---
title: "Simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
have_tidy <- requireNamespace("dplyr", quietly = TRUE)&&
  requireNamespace("ggplot2", quietly = TRUE)

if(!have_tidy){
  message("To reproduce the full vignette, please install the suggested packages: dplyr, ggplot2")
}
```

```{r setup}
library(dplyr)
library(ggplot2)
library(simsurv)
library(patchwork)
#library(bhCRR)
```

# Motivation

Outline a customizable scenario builder so that we can pick and choose which pieces are important.

Every choice that I originally made

-   Design matrix based off of a true dataset

    -   Fixed n, p, distribution (non-zero values, comes from low abundance CLR data)

-   Permutations (for low correlation)

-   Number of competing risks

-   Number of true covariates in each set

    -   Fixed sparsity, masking

-   Effect sizes

    -   Range log HR distribution

    -   Target standard deviation of linear predictors

    -   Masked variables as a multiplicative factor

-   Outcome parametric distribution

    -   Linear form

    -   Weibull ph

        -   Fixed shape and scale

# Specifications

## Helper functions

```{r}

`%||%` <- function(a, b) if (!is.null(a)) a else b

sim_merge_lists <- function(default, user) {
  if (is.null(user)) return(default)
  if (!is.list(user)) stop("Expected a list, got: ", class(user)[1])
  out <- default
  for (nm in names(user)) {
    if (!nm %in% names(out)) {
      out[[nm]] <- user[[nm]]
    } else if (is.list(out[[nm]]) && is.list(user[[nm]])) {
      out[[nm]] <- sim_merge_lists(out[[nm]], user[[nm]])
    } else {
      out[[nm]] <- user[[nm]]
    }
  }
  out
}

```

## List of specification

Things to add:

-   Nonlinear and interaction terms in the truth

```{r}
sim_spec <- function(
  n = 50, # samples
  p = 100, # features
  x = list(), # design matrix
  truth = list(), # variable selection
  risks = list(), # risk 
  censor = list(), #type of censoring
  seed = NULL
  
){
  
  
defaults <- list(
    n = n,
    p = p,
    x = list(
      family = "gaussian",  #  gaussian | binary_latent | zip | mixed
      corr = list(type = "indep", rho = 0.0),  # indep | ar1 | block | factor
      zero_inflation = 0.0,
      standardize = TRUE,

      # optional knobs used by some families/structures
      binary_prob = 0.5,      # for binary_latent
      zip_rate = 1.0,         # for zip (rough first pass)
      factor_k = 5            # for corr$type = "factor"
    ),
    truth = list(
      share_active_set = 0,
      shared_multiplier = 1,
      sparsity = 0.05,
      beta = list(dist = "normal", mean = 0, sd = c(0.3, 0.3)),

      # nonlinear = list(
      #   enabled = FALSE,
      #   n_terms = 0,
      #   forms = c("sin", "square"),
      #   sd = 0.3
      # ),
      # interactions = list(
      #   enabled = FALSE,
      #   n_pairs = 0,
      #   sd = 0.2
      # )
      
      target_eta_sd = c(0.5, 0.5)
    ),
    risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.011),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.011),
        link = "PH"
      )
      # cluster = list(
      #   enabled = FALSE,
      #   G = 40,
      #   sd = 0.35,
      #   shared_across_causes = TRUE
      # )
    ),
    censor = list(
      type = "administrative", # none | administrative | random | administrative_plus_random
      administrative_time = 65,
      random = list(dist = "exponential", rate = 0.15)
      # future: target_censor, calibrate, etc.
    ),
    seed = seed
  )

  spec <- defaults
  spec$x <- sim_merge_lists(defaults$x, x)
  spec$truth <- sim_merge_lists(defaults$truth, truth)
  spec$risks <- sim_merge_lists(defaults$risks, risks)
  spec$censor <- sim_merge_lists(defaults$censor, censor)

  class(spec) <- "sim_spec"

  spec
}
```

## Generator wrapper

```{r}
sim_generate <- function(spec,
                         custom_design_matrix = NULL,
                         custom_predictor_set = NULL,
                         custom_betas = NULL,
                         seed = NULL){
  if (!inherits(spec, "sim_spec")) {
    # allow passing a plain list as long as it has the right shape
    class(spec) <- unique(c("sim_spec", class(spec)))
  }
  
  if (!is.null(seed)) {
    set.seed(seed)
  } else if (!is.null(spec$seed)) {
    set.seed(spec$seed)
  }
  
  n <- as.integer(spec$n)
  p <- as.integer(spec$p)
  
  # Design matrix "X"
  ## Allow for custom design matrix
  if(!is.null(custom_design_matrix)){
    X <- data.matrix(custom_design_matrix)
  }else{
    
    X <- sim_generate_x(spec$x, n, p)
  }
  
  print("Success in generating X")
  
  truth <- spec$truth
  
  # Variable selection "truth:
  ## Allow for custom covariate choices, but check that custom predictor set is a list of length 2 
  if(!is.null(custom_predictor_set)){
     if(!(typeof(custom_predictor_set) == "list" & 
                length(custom_predictor_set) == 2 &
                sum(custom_predictor_set[[1]]%%1) == 0 &
                sum(custom_predictor_set[[2]]%%1) == 0 &
                max(custom_predictor_set[[1]]) <= p &
                max(custom_predictor_set[[2]]) <= p)){
    warning("custom_predictor_set should be a list of 2 sets of index integers (with max integer <= p). Ignoring custom predictor sets and custom_betas. Generating from truth defaults. ")
    custom_predictor_set <- NULL
    custom_betas <- NULL
    }}
  
    
  if(!is.null(custom_predictor_set)){
    truth$shared_multiplier = spec$truth$shared_multiplier
    truth$cause1 = custom_predictor_set[[1]]
    truth$cause2 = custom_predictor_set[[2]]
    truth$shared_predictors = intersect(custom_predictor_set[[1]],
                                        custom_predictor_set[[2]])
    
    #print real truth rather than defaults
    #number of real shared active sets
    truth$share_active_set = length(truth$shared_predictors)
    #true sparsity based on min sparsity of the 2 causes
    truth$sparsity = min(c(length(truth$cause1), length(truth$cause1)))/p
    #placeholders
    truth$beta = spec$truth$beta
    truth$target_eta_sd = spec$truth$target_eta_sd
  }else{
    truth <- sim_generate_truth(spec$truth, p, seed = seed)
  }
  
  print("Success in generating truth")
  
  
  # Covariate effects (add to "truth")
  ## Allow for custom covariate effects, but check that custom predictors is a list of 
  if(!is.null(custom_betas)){
     if(!(length(custom_betas[[1]]) == length(truth$cause1) &
       length(custom_betas[[2]]) == length(truth$cause2))){
    warning("Lengh of custom betas does not match length of predictor sets. Generating from truth defaults.")
    custom_betas <- NULL
  }}
  ## Covariate effects and linear predictors
  if(!is.null(custom_betas)){
    
    effects1 <- list()
    effects2 <- list() 
    
    
    effects1$betas1 <- custom_betas[[1]]
    effects1$betas1_sd <- sd(effects1$betas1)
    effects1$eta1 <- X[, truth$cause1] %*% custom_betas[[1]]
    effects1$eta1_mean = mean(effects1$eta1)
    effects1$eta1_sd <- sd(effects1$eta1)
    
    effects2$betas2 <- custom_betas[[2]]
    effects2$betas2_sd <- sd(effects2$betas2)
    effects2$eta2 <- X[, truth$cause2] %*% custom_betas[[2]]
    effects2$eta2_mean <- mean(effects2$eta2)
    effects2$eta2_sd <- sd(effects2$eta2)
    
    effects <- list(effects1 = effects1, effects2 = effects2)
    
    
  }else{
  effects <- sim_compute_effects(X, truth)
  }
  
  if(truth$share_active_set > 0){
      #sort to the shared predictor order
      i1 <- match(truth$shared_predictors, truth$cause1)
      i2 <- match(truth$shared_predictors, truth$cause2)
    truth$shared_multiplier <- effects$effects1$betas1[i1] / effects$effects2$betas2[i2]
    }else{
      truth$shared_multiplier <- NA
    }
  

  truth$beta <- NULL
  truth$target_eta_sd <- NULL
  truth$effects <- effects
  
  #print("Shared predictors")
  #print(truth$shared_predictors)
  
  print("Success in generating truth with outcomes")
  #test_truth <<- truth

  # Outcome "risks"
  ## baseline hazard
  b1 <- spec$risks$cause1$baseline
  b2 <- spec$risks$cause2$baseline

  ## Latent competing event times
  event1 <- sim_weibull_ph(shape = b1$shape %||% 1, 
                       scale = b1$scale %||% 0.011, 
                       eta_val = truth$effects$effects1$eta1)
  
  T1 <- event1$eventtime
  
  event2 <- sim_weibull_ph(shape = b2$shape %||% 1, 
                       scale = b2$scale %||% 0.011, 
                       eta_val = truth$effects$effects2$eta2)
  T2 <- event2$eventtime
  
  print("Success in generating latent event times")

  maxtime <- max(c(T1, T2))
  # Censoring "censor"
  C <- sim_generate_censor_time(spec$censor, n, maxtime)
  #print(C)
  time <- pmin(T1, T2, C)

  event1 <- (T1 <= T2) & (T1 <= C)
  event2 <- (T2 <  T1) & (T2 <= C)
  status <- integer(n)
  status[event1] <- 1L
  status[event2] <- 2L
  # remaining are censored (0)
  
  print("Success in generating censoring")

  out <- list(
    id = seq_len(n),
    time = as.numeric(time),
    status = as.integer(status),
    latent_time_event1 = T1,
    latent_time_event2 = T2,
    latent_time_censor = C,#ifelse(spec$censor$type == "none", NA, C),
    X = X,
    spec_call = spec,
    truth = truth
  )


  class(out) <- "sim_data"
  out
  
  
}
```

# Design Matrix

-   For future implementation. ChatGPT 5.2 Pro placeholder.

-   Need to add clinical factors as a designation thing

```{r}
sim_generate_x <- function(xspec, n, p) {
  corr_type <- xspec$corr$type %||% "indep"
  rho <- xspec$corr$rho %||% 0

  # 1) latent Gaussian core with requested dependence
  Z <- switch(
    corr_type,
    indep = matrix(stats::rnorm(n * p), nrow = n),
    ar1 = {
      if (!is.finite(rho) || abs(rho) >= 1) stop("For ar1, corr$rho must satisfy abs(rho) < 1.")
      eps <- matrix(stats::rnorm(n * p), nrow = n)
      out <- eps
      if (p >= 2) {
        s <- sqrt(1 - rho^2)
        for (j in 2:p) out[, j] <- rho * out[, j - 1] + s * eps[, j]
      }
      out
    },
    block = {
      block_size <- xspec$corr$block_size %||% 50L
      block_size <- as.integer(block_size)
      if (block_size < 1) stop("For block correlation, corr$block_size must be >= 1.")
      # equicorrelation within blocks via a shared latent factor
      out <- matrix(0, nrow = n, ncol = p)
      nb <- ceiling(p / block_size)
      for (b in seq_len(nb)) {
        j0 <- (b - 1) * block_size + 1
        j1 <- min(b * block_size, p)
        m <- j1 - j0 + 1
        z_shared <- stats::rnorm(n)
        eps <- matrix(stats::rnorm(n * m), nrow = n)
        out[, j0:j1] <- sqrt(max(rho, 0)) * z_shared + sqrt(max(1 - max(rho, 0), 0)) * eps
      }
      out
    },
    factor = {
      k <- as.integer(xspec$factor_k %||% 5L)
      k <- max(1L, k)
      Zf <- matrix(stats::rnorm(n * k), nrow = n)
      Lambda <- matrix(stats::rnorm(k * p), nrow = k) / sqrt(k)
      E <- matrix(stats::rnorm(n * p), nrow = n)
      Zf %*% Lambda + E
    },
    stop("Unknown corr$type: ", corr_type)
  )

  # 2) transform latent Z to requested family
  family <- xspec$family %||% "gaussian"
  X <- switch(
    family,
    gaussian = Z,
    binary_latent = {
      prob <- xspec$binary_prob %||% 0.5
      prob <- min(max(prob, 1e-6), 1 - 1e-6)
      thr <- stats::qnorm(1 - prob)
      (Z > thr) * 1.0
    },
    zip = {
      # Rough first pass: correlated log-intensity from Z
      rate0 <- xspec$zip_rate %||% 1.0
      mu <- log(pmax(rate0, 1e-8)) + 0.3 * Z
      stats::rpois(n * p, lambda = exp(mu)) |> matrix(nrow = n)
    },
    mixed = {
      # first pass: 70% gaussian, 30% binary_latent
      p_bin <- as.integer(round(0.3 * p))
      idx_bin <- if (p_bin > 0) sample.int(p, p_bin) else integer(0)
      Xmix <- Z
      if (length(idx_bin) > 0) Xmix[, idx_bin] <- (Z[, idx_bin] > 0) * 1.0
      Xmix
    },
    stop("Unknown x$family: ", family)
  )

  # 3) zero inflation overlay (applies to all families in this first pass)
  zi <- xspec$zero_inflation %||% 0
  if (zi > 0) {
    mask <- matrix(stats::runif(n * p) < zi, nrow = n)
    X[mask] <- 0
  }

  # 4) optional standardization
  if (isTRUE(xspec$standardize)) {
    cm <- colMeans(X)
    cs <- apply(X, 2, stats::sd)
    cs[cs == 0] <- 1
    X <- sweep(X, 2, cm, "-")
    X <- sweep(X, 2, cs, "/")
  }

  X
}
```

# Working Model

Things to add

-   Different number of predictors per outcome event

-   Random vs preset predictors

-   Clinical predictors

-   Different distributions of the betas

```{r}
sim_generate_truth <- function(trspec, p, seed){
  
  # Number of active predictors
  if(trspec$sparsity < 1){
    s <- ceiling(p * (trspec$sparsity %||% 0.05))
    s <- max(0L, min(p, as.integer(s)))
  }else if(trspec$sparsity >= 1 & trspec$sparsity <= p ){
    s <- trspec$sparsity
  }
  
  #print(s)
  
  # Choice of active predictors
  if(!is.null(seed)){
    set.seed(seed)
  }else if(is.null(seed)){
    set.seed(9134)
  }
  active_predictors <- sample(1:p, 2*s-trspec$share_active_set)
  active_predictors_cause1 <- active_predictors[1:s]
  active_predictors_cause2 <- active_predictors[((s-trspec$share_active_set)+1):length(active_predictors)]
  shared_predictors <- dplyr::intersect(active_predictors_cause1,
                                        active_predictors_cause2)
  
  #rm(.Random.seed)
  
  ret <- list(shared_multiplier = trspec$shared_multiplier,
              cause1 = active_predictors_cause1,
              cause2 = active_predictors_cause2,
              shared_predictors = shared_predictors,
              share_active_set = trspec$share_active_set,
              sparsity = trspec$sparsity,
              target_eta_sd = trspec$target_eta_sd,
              beta = trspec$beta)
  
  return(ret)
  
}
```

```{r}
sim_compute_effects <- function(X, truth) {
  
  # 1. Mean-center the entire design matrix upfront
  # This guarantees all future eta = X %*% beta have a mean of 0
  X <- scale(X, center = TRUE, scale = FALSE)
  
  # Simulate event of interest first
  X1 <- X[, truth$cause1, drop = FALSE]
  
  target_beta_sd <- truth$beta$sd[1]
  target_eta_sd <- truth$target_eta_sd[1]
  p1 <- ncol(X1)
  
  # Generate and scale betas for Cause 1
  betas1 <- rnorm(p1, 0, target_beta_sd)
  eta1_0 <- as.numeric(X1 %*% betas1)
  print(mean(eta1_0))
  
  # Scale beta to hit the exact target SD for eta
  scale_factor <- target_eta_sd / sd(eta1_0)
  betas1 <- betas1 * scale_factor
  
  # eta1 now naturally has a mean of 0 and the exact target SD
  eta1 <- as.numeric(X1 %*% betas1)
  print(mean(eta1))
  
  # Simulate competing event next
  X2 <- X[, truth$cause2, drop = FALSE]
  
  if (length(truth$beta$sd) >= 2) {
    target_beta_sd <- truth$beta$sd[2]
  }
  if (length(truth$target_eta_sd) >= 2) {
    target_eta_sd <- truth$target_eta_sd[2]
  }
  
  p2 <- ncol(X2)
  betas2 <- numeric(p2)
  
  if (truth$share_active_set < 1) {
    betas2 <- rnorm(p2, 0, target_beta_sd)
    eta2_0 <- as.numeric(X2 %*% betas2)
    scale_factor <- target_eta_sd / sd(eta2_0)
    betas2 <- betas2 * scale_factor
  } else {
    i1 <- match(truth$shared_predictors, truth$cause1)
    i2 <- match(truth$shared_predictors, truth$cause2)
    
    if (length(truth$shared_multiplier) == 1) {
      truth$shared_multiplier <- rep(truth$shared_multiplier, truth$share_active_set)
    }
    
    betas2[i2] <- betas1[i1] * (1 / truth$shared_multiplier)
    betas2[-i2] <- rnorm(p2 - length(i2), 0, target_beta_sd)
    
    eta_overlap <- as.numeric(X2[, i2, drop = FALSE] %*% betas2[i2])
    eta_nonoverlap_raw <- as.numeric(X2[, -i2, drop = FALSE] %*% betas2[-i2])
    
    # Use optimization to solve for the correct scaling factor
    objective <- function(c) {
      (sd(eta_overlap + c * eta_nonoverlap_raw) - target_eta_sd)^2
    }
    
    opt_res <- optimize(objective, interval = c(-100, 100))
    betas2[-i2] <- opt_res$minimum * betas2[-i2]
  }
  
  # eta2 also naturally has a mean of 0 and the exact target SD
  eta2 <- as.numeric(X2 %*% betas2)
  
  ret <- list(
    effects1 = list(
      betas1 = betas1,
      betas1_sd = sd(betas1),
      eta1 = eta1,
      eta1_mean = mean(eta1),
      eta1_sd = sd(eta1)
    ),
    effects2 = list(
      betas2 = betas2,
      betas2_sd = sd(betas2),
      eta2 = eta2,
      eta2_mean = mean(eta2),
      eta2_sd = sd(eta2)
    )
  )
  
  return(ret)
}
```

# Latent Outcome Model

## Weibull proportional hazards model

Parameterization of the weibull model based off of the weibull distribution with shape parameter $\gamma > 0$ and scale/rate parameter $\lambda > 0$ . Shape controls whether the hazard increases or decreases with time. As $\lambda$ increases, the hazard increases, and survival is shorter. Note the parameterization is different here than in parameterizations which relate scale $b$ to xx, which can be achieved with the transformation $b = \lambda ^{-1/\gamma}$.

Weibull proportional hazards relationships.

Baseline hazard function

$$
h_0(t) = \gamma\lambda t^{\gamma-1}
$$

Baseline cumulative hazard function

$$
H_0(t) = \int_0^t h_0(s) ds \\
=  \lambda \int_0^t \gamma s^{\gamma-1} ds\\
= \lambda t^{\gamma}\\
= -\log(S_0(t))
$$

Baseline survival function

$$
S_0(t) = \exp(-H_0(t)) \\ 
= \exp(-\lambda t^{\gamma})\\
= 1 - F_0(t)
$$

Baseline probability density function

$$
f_0(t) = S_0(t) h_0(t)\\
=  \exp(-\lambda t^{\gamma})\gamma\lambda t^{\gamma-1}\\
= -\frac{d}{dt}S_0(t)\\
= \frac{d}{dt}F_0(t)
$$

Baseline cumulative incidence function

$$
F_0(t) = 1- S_0(t) \\
= 1 - \exp(-\lambda t^{\gamma})
$$

This cumulative incidence function is a weibull CDF.

With linear predictors given by $\eta_i = x_i^T \beta$, the person specific functions are

Person specific hazard function

$$
h_i(t) = h_0(t) \exp(\eta_i) \\
=\gamma\lambda t^{\gamma-1} \exp(\eta_i)
$$

Person specific cumulative hazard function

$$
H_i(t) = \int_0^t h_i(s) ds \\
=  \lambda \exp(\eta_i) \int_0^t \gamma s^{\gamma-1} ds\\
= \lambda t^{\gamma} \exp(\eta_i)\\
= -\log(S_it))
$$

Person specific survival function

$$
S_i(t) = \exp(-H_i(t)) \\ 
= \exp(-\lambda t^{\gamma}\exp(\eta_i))\\
= 1 - F_i(t)
$$

Person specific density function

$$
f_i(t) = S_i(t) h_i(t)\\
=  \exp(-\lambda t^{\gamma}\exp(\eta_i))\gamma\lambda t^{\gamma-1}\exp(\eta_i)\\
= -\frac{d}{dt}S_i(t)\\
= \frac{d}{dt}F_i(t)
$$

Person specific cumulative distribution function

$$
F_i(t) = 1- S_i(t) \\
= 1 - \exp(-\lambda t^{\gamma}\exp(\eta_i))
$$

The effective scale of the

### Randomness of Eta

Consider $\eta \sim N(0, \sigma^2)$ to be a random pull. Then $\exp(\eta) \sim$ lognormal. This means that only the conditional event time $T_i | \eta_i$ is weibull ditributed.

## Default values

Generic value of gamma: gamma = 1

-   Constant hazard: 1

-   Moderately increasing hazard: [1.2, 1]

-   Decreasing hazard due to early risk: \< 1

Generic values of lambda

-   Choose a median survival time $m$ so that we can solve for

    $$
    m:= arg{ S_0(m) = 0.5}\\
     \exp(-\lambda m^{\gamma}) = 0.5 \\
     \lambda = -\frac{log 0.5}{m ^{\gamma}}
    $$

-   Choose a baseline event probability $p$ at a certain time $\tau$

    $$
    p:= Pr(T_{event} \leq \tau) = 1 - S_0(\tau) \\
    S_0(\tau) = 1-p\\
    \exp(-\lambda \tau ^ {\gamma}) = 1-p \\
    \lambda = \frac{-\log(1-p)}{\tau^{\gamma}}
    $$

Generic values of sd of eta

-   Mild: 0.3 (HR increases 1.35 per standard deviation)

## Function

```{r}
sim_weibull_ph <- function(shape, scale, eta_val){
  x_frame <- data.frame(eta = eta_val)


  names(x_frame)<- "eta"
  betas_frame <- c(eta = 1)

  simsurv::simsurv(dist = "weibull",
                   lambdas = scale,
                   gammas = shape,
                   x =x_frame,
                   betas = betas_frame)
}
```

# Censoring

```{r}
sim_generate_censor_time <- function(cspec, n, maxtime){
  if(cspec$type == "none") {
    return(rep(maxtime + 10, n))
  }else if(cspec$type == "administrative") {
    return(rep(cspec$administrative_time, n))
  }
  
  random_censor_dist <- cspec$random$dist
  random_censor_admin <- cspec$administrative_time
  latent_censor_time <- switch(
    random_censor_dist,
    exponential = {
      temp <- rexp(n, rate = cspec$random$rate)
      temp[temp>=cspec$administrative_time]<- cspec$administrative_time
      temp
      },
    #exponential ={rexp(n, rate = 1/(random_censor_admin+30))},
    stop("Unknown random censor distribution: ", random_censor_dist)
  )
  return(latent_censor_time)
}
```

# Getting defaults

```{r}
fpath <- system.file( "example_dat.Rdata", package = "bhCRR")

if (nzchar(fpath)) {
  load(fpath)
  # The objects are now in your environment
} else {
  warning("Could not find the dataset in the installed package. Loading from inst file.")
  
  load("../inst/example_dat.Rdata") 
}

permute_rows <- function(df, seed = NULL) {
  if (!is.null(seed)) set.seed(seed)
  m <- as.matrix(df)                # numeric matrix
  permuted_mat <- t(apply(m, 1, function(row) row[sample.int(length(row))]))
  colnames(permuted_mat) <- colnames(m)
  rownames(permuted_mat) <- rownames(m)
  as.data.frame(permuted_mat)
}

example_dm <- example_dat |> dplyr::select(dplyr::starts_with("ASV")) |>
  permute_rows()
```

```{r}
scenario_default <- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
  truth = list(share_active_set = 2,
               shared_multiplier = c(0.5, 0.2),
               sparsity = 0.05,
               beta = list(dist = "normal", mean = 0, sd = c(0.3, 0.3)),
               target_eta_sd = c(0.5, 0.5)), # variable selection
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.011),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.011),
        link = "PH"
      )), # risk 
  censor = list(type = "administrative",
                administrative_time = 65), #type of censoring
  seed = 9134
  
)


```

```{r}
scenario_default_sim <- sim_generate(scenario_default,
                         custom_design_matrix = example_dm,
                         custom_predictor_set = NULL,
                         custom_betas = NULL,
                         seed = 9134)
```

# Prelim Diagnostics

To show that we are successfully able to control outcomes, I will generate a scenarios based on a specified shape=1 (constant hazard) and scale values (calculated for baseline median survival time of 50 days) for the weibull proportional hazards distribution. I will generate 10 sets of this scenario to check whether we are 1) Getting the expected weibull distribution for the latent event times and 2) Getting the desired event 1/event 2/censoring proportion.

## Checking generated default parameters

I will fix the design matrix, the predictor set for both event 1 (of interest) and event 2 (competing risk), and the covariate effect for both events based on the default from scenario_default_sim (run above).

The important predictors for event 1.

```{r}
names(example_dm[,scenario_default_sim$truth$cause1])
```

The betas and corresponding hazard ratios.

```{r}
scenario_default_sim$truth$effects$effects1$betas1
exp(scenario_default_sim$truth$effects$effects1$betas1)
```

The distribution of the linear predictors. The standard deviation should be 0.5 as specified by the default.

```{r}
lp_event1 <-scale(as.matrix(example_dm[,scenario_default_sim$truth$cause1]), 
      center = TRUE, scale = FALSE) %*% 
  scenario_default_sim$truth$effects$effects1$betas1

mean(lp_event1)
sd(lp_event1)
```

```{r}
hist(lp_event1)
```

The important predictors for event 2.

```{r}
names(example_dm[,scenario_default_sim$truth$cause2])
```

The betas and corresponding hazard ratios.

```{r}
scenario_default_sim$truth$effects$effects2$betas2
exp(scenario_default_sim$truth$effects$effects2$betas2)
```

The distribution of the linear predictors. The standard deviation should be near 0.5 as specified by the default, but currently trouble shooting on how to target the sd while inducing masking effects. (See commented out section on objective solver in sim_compute_effects).

```{r}

lp_event2 <-scale(as.matrix(example_dm[,scenario_default_sim$truth$cause2]), 
      center = TRUE, scale = FALSE) %*% 
  scenario_default_sim$truth$effects$effects2$betas2


mean(lp_event2)
sd(lp_event2)

```

```{r}
hist(lp_event2)
```

## Defining basic scenarios for testing

Let us consider 3 different scenarios.

1) Basic scenario no censoring: Shape parameter for both events is 1 (constant hazard). This is equivalent to the exponential distribution. We will have no censoring. And we will have the scale parameter be the same for the 2 events. This should produce about equal event rates. We can choose the scale values based on a median survival time of 50 days.

2) Basic scenario with administrative censoring: We will keep the same set up of the basic scenario but add in administrative censoring at 90 days.

3) Basic scenario with administrative and random censoring: We will keep the same set up of the basic scenario but add in administrative censoring at 90 days and random censoring with an exponential distribution with rate such that prior to the administrative censoring there is a 15% chance of being censored.

## Choosing scales

If $\gamma = 1$ and we want median survival time to be 50, then $\lambda = -log(0.5)/50^1 = 0.0139$. This would correspond to a event probability at 50 days of $1-\exp(-0.0139 * 50^1)=0.5$ and an event probability at 90 day (the time of administrative censoring) to be 0.71. For the random censoring, let's say that we would want random censoring to occur at a rate near 15% (if there were no other events). Then we can solve the rate parameter (for the exponential distribution) by $\lambda_{cen} = -\log(1- 0.15)/90= 0.0018$.

## Diagnostic plots

```{r}
outcome_simulation_diagnostic_plots <- function(sim, shapes, scales){
  df_event1 <- cbind(  
    data.frame(id = sim$id, time = sim$latent_time_event1, status = 1),
    scale(sim$X, center = TRUE, scale = FALSE)[,sim$truth$cause1 ]
  )
  
  
  df_event2 <- cbind(  
    data.frame(id = sim$id, time = sim$latent_time_event2, status = 1),
    scale(sim$X, center = TRUE, scale = FALSE)[,sim$truth$cause2 ]
  )
  
  
  if(shapes[1] == shapes[2] &
     scales[1] == scales[2]){
    
    plot1 <- ggplot(df_event1, aes(x = time)) +
      # Map color inside aes() so scale_color_manual works
      geom_density(aes(color = "Event 1 (latent)"), linewidth = 1) +
      geom_density(data = df_event2, aes(color = "Event 2 (latent)"), linewidth = 1) +
      
      # Theoretical Curves
      stat_function(fun = dweibull, 
                    args = list(shape = shapes[1], scale = scales[1]^(-1)),
                    aes(color = "Theoretical"), 
                    linewidth = 1) +
      
      
      # Manual Color Scale
      scale_color_manual(
        name = "Distribution",
        values = c(
          "Event 1 (latent)" = "darkred", 
          "Event 2 (latent)" = "darkblue", 
          "Theoretical" = "black"
        )
      ) +
      labs(
        title = "Observed vs. Theoretical Marginal Event Density",
        x = "Time",
        y = "Density"
      ) +
      theme_classic()
    
  }else{
    plot1 <- ggplot(df_event1, aes(x = time)) +
      # Map color inside aes() so scale_color_manual works
      geom_density(aes(color = "Event 1 (latent)"), linewidth = 1) +
      geom_density(data = df_event2, aes(color = "Event 2 (latent)"), linewidth = 1) +
      
      # Theoretical Curves
      stat_function(fun = dweibull, 
                    args = list(shape = shapes[1], scale = scales[1]^(-1)),
                    aes(color = "Event 1 (theoretical)"), 
                    linewidth = 1) +
      stat_function(fun = dweibull, 
                    args = list(shape = shapes[2], scale = scales[2]^(-1)),
                    aes(color = "Event 2 (theoretical)"), 
                    linewidth = 1) +
      
      # Manual Color Scale
      scale_color_manual(
        name = "Distribution",
        values = c(
          "Event 1 (latent)" = "darkred", 
          "Event 2 (latent)" = "darkblue", 
          "Event 1 (theoretical)" = "red", 
          "Event 2 (theoretical)" = "blue"
        )
      ) +
      labs(
        title = "Observed vs. Theoretical Marginal Event Density",
        x = "Time",
        y = "Event Probability Density"
      ) +
      theme_classic()
  }
  
  plot1
  
  cox_snell_func <- function(df){
    form1 <- names(df %>%
                     select(-c(id, time, status))) %>%
      paste(collapse = " + ")


    form2 <- paste0("Surv(time, status) ~ ", form1)

    cox_model <- coxph(as.formula(form2), data = df)


    cs_residual_fit <- df$status - residuals(cox_model, type = "martingale")

    fit_cs_emp <- survfit(Surv(cs_residual_fit, df$status) ~ 1)
    cum_haz_emp <- -log(fit_cs_emp$surv)

    cox_snell_df <- data.frame(cs_residuals = fit_cs_emp$time,
                               cumhaz = cum_haz_emp)

    ggplot(cox_snell_df, aes(x = cs_residuals, y = cumhaz))+
      geom_step(linewidth = 0.8)+
      geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed", linewidth = 1) +
      labs(
        # title = "Cox-Snell Residuals Diagnostic Plot",
        x = "Cox-Snell Residuals",
        y = "Estimated Cumulative Hazard"
      ) +
      theme_classic()
  }

  plot2a <- cox_snell_func(df_event1)+
    ggtitle("Event 1")

  plot2b <- cox_snell_func(df_event2)+
    ggtitle("Event 2")

  plot2 <- plot2a + plot2b +
    plot_annotation(title = "Cox-Snell Residual Diagnostic Plots of Latent Event Times")

  plot2
  
  table1 <- data.frame(time = sim$time, status = sim$status)%>% 
    mutate(status = factor(status, 
                           levels = c(0,1,2), 
                           labels = c("Censored", "Event1", "Event2"))) %>%
    group_by(status, .drop = F)%>% 
    summarize(N = n(), 
              Observed_TTE_median = median(time), 
              Observed_TTE_mean = mean(time), 
              Observed_TTE_sd = sd(time))%>% 
    mutate(Proportion = N/sum(.$N))%>%
    select(status, N, Proportion, Observed_TTE_median, Observed_TTE_mean, Observed_TTE_sd)%>%
    ungroup()%>%
    mutate(Latent_TTE_median = case_when(status == "Censored" ~ median(sim$latent_time_censor, na.rm = T),
                                         status == "Event1" ~ median(sim$latent_time_event1, na.rm = T),
                                         status == "Event2" ~ median(sim$latent_time_event2, na.rm = T)))%>%
    mutate(Latent_TTE_mean = case_when(status == "Censored" ~ mean(sim$latent_time_censor, na.rm = T),
                                         status == "Event1" ~ mean(sim$latent_time_event1, na.rm = T),
                                         status == "Event2" ~ mean(sim$latent_time_event2, na.rm = T)))%>%
    mutate(Latent_TTE_sd = case_when(status == "Censored" ~ sd(sim$latent_time_censor, na.rm = T),
                                         status == "Event1" ~ sd(sim$latent_time_event1, na.rm = T),
                                         status == "Event2" ~ sd(sim$latent_time_event2, na.rm = T)))
  
  ret <- list("Marginal" = plot1,
              "CoxSnell" = plot2,
              "ObservedOutcomeSummary" = table1)
  
  
  return(ret)
  
}
```

# Basic scenarios

## Basic scenario, no censoring

```{r}
basic_NoCensoring<- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
   truth = list(#share_active_set = 2,
                shared_multiplier = c(0.5, 0.2)#,
                #sparsity = 0.05,
                #beta = list(dist = "normal", mean = 0, sd = c(0.3, 0.3)),
                #target_eta_sd = c(0.5, 0.5)), # variable selection
   ),
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      )), # risk 
  censor = list(type = "none",
                administrative_time = 65), #type of censoring
  seed = NULL
  
)

tenSeeds <- .Random.seed[1:10]
basic_NoCensoring_sim <- list()

for(i in 1:10){
basic_NoCensoring_sim[[i]] <- sim_generate(basic_NoCensoring,
                         custom_design_matrix = scale(example_dm,
                                                      center = TRUE,
                                                      scale = FALSE),
                         custom_predictor_set = list(scenario_default_sim$truth$cause1,
                                                  scenario_default_sim$truth$cause2),
                         custom_betas = 
                list(scenario_default_sim$truth$effects$effects1$betas1,
                     scenario_default_sim$truth$effects$effects2$betas2),
                         seed = tenSeeds[i])
}
```

### Checking event times

Median event times should be around 50 days (variability from the standard deviation of the eta).

```{r}
outcome_simulation_diagnostic_plots(basic_NoCensoring_sim[[2]], shapes = c(1,1), scales = c(0.0139,0.0139))
```

```{r}

summary_tables <- list()
for(i in 1:10){
temp1<-outcome_simulation_diagnostic_plots(basic_NoCensoring_sim[[i]], shapes = c(1,1), scales = c(0.0139,0.0139))
temp2<- temp1$ObservedOutcomeSummary
temp2$SimNumber = rep(i, nrow(temp2))
summary_tables[[i]]<- temp2
rm(temp2)
}
basic_NoCensoring_Summary = do.call("rbind", summary_tables)

basic_NoCensoring_Summary

```

```{r}
basic_NoCensoring_Summary %>%
  group_by(status)%>%
  summarise(Mean_Proportion = mean(Proportion, na.rm = T),
            SD_Proportion = sd(Proportion, na.rm = T))
```

```{r}
basic_NoCensoring_Summary %>%
  select(status, Observed_TTE_median, Latent_TTE_median)%>%
  tidyr::pivot_longer(cols = c(Observed_TTE_median, Latent_TTE_median),
                      names_to = "True",
                      values_to = "TTE_Median")%>%
  ungroup()%>%
  filter(status != "Censored")%>%
  ggplot(aes(x = status, y = TTE_Median, color = True))+
  geom_boxplot()+
  ylab("Median Time to Event")+
  xlab("Outcome")+
  scale_color_manual(name = "",
                     values = c("red", "blue"),
                     labels = c("Latent", "Observed"))+
  theme_classic()+
  ggtitle("Observed and Latent Event Times Across 10 Runs")
```

## Basic scenario, admin censoring

```{r}
basic_AdminCensoring<- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
   truth = list(#share_active_set = 2,
                shared_multiplier = c(0.5, 0.2)#,
                #sparsity = 0.05,
                #beta = list(dist = "normal", mean = 0, sd = c(0.3, 0.3)),
                #target_eta_sd = c(0.5, 0.5)), # variable selection
   ),
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      )), # risk 
  censor = list(type = "administrative",
                administrative_time = 90), #type of censoring
  seed = NULL
  
)

#tenSeeds <- .Random.seed[1:10]
basic_AdminCensoring_sim <- list()

for(i in 1:10){
basic_AdminCensoring_sim[[i]] <- sim_generate(basic_AdminCensoring,
                         custom_design_matrix = scale(example_dm,
                                                      center = TRUE,
                                                      scale = FALSE),
                         custom_predictor_set = list(scenario_default_sim$truth$cause1,
                                                  scenario_default_sim$truth$cause2),
                         custom_betas = 
                list(scenario_default_sim$truth$effects$effects1$betas1,
                     scenario_default_sim$truth$effects$effects2$betas2),
                         seed = tenSeeds[i])
}
```

```{r}


outcome_simulation_diagnostic_plots(basic_AdminCensoring_sim[[2]], shapes = c(1,1), scales = c(0.0139,0.0139))
```

```{r}
summary_tables <- list()
for(i in 1:10){
temp1<-outcome_simulation_diagnostic_plots(basic_AdminCensoring_sim[[i]], shapes = c(1,1), scales = c(0.0139,0.0139))
temp2<- temp1$ObservedOutcomeSummary
temp2$SimNumber = rep(i, nrow(temp2))
summary_tables[[i]]<- temp2
rm(temp2)
}
basic_AdminCensoring_Summary = do.call("rbind", summary_tables)

basic_AdminCensoring_Summary
```

```{r}
basic_AdminCensoring_Summary %>%
  group_by(status)%>%
  summarise(Mean_Proportion = mean(Proportion, na.rm = T),
            SD_Proportion = sd(Proportion, na.rm = T))
```

```{r}
basic_AdminCensoring_Summary %>%
  select(status, Observed_TTE_median, Latent_TTE_median)%>%
  tidyr::pivot_longer(cols = c(Observed_TTE_median, Latent_TTE_median),
                      names_to = "True",
                      values_to = "TTE_Median")%>%
  ungroup()%>%
  ggplot(aes(x = status, y = TTE_Median, color = True))+
  geom_boxplot()+
  ylab("Median Time to Event")+
  xlab("Outcome")+
  scale_color_manual(name = "",
                     values = c("red", "blue"),
                     labels = c("Latent", "Observed"))+
  theme_classic()+
  ggtitle("Observed and Latent Event Times Across 10 Runs")
```

## Basic scenario, random censoring

```{r}
basic_RandomCensoring<- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
   truth = list(#share_active_set = 2,
                shared_multiplier = c(0.5, 0.2)#,
                #sparsity = 0.05,
                #beta = list(dist = "normal", mean = 0, sd = c(0.3, 0.3)),
                #target_eta_sd = c(0.5, 0.5)), # variable selection
   ),
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      )), # risk 
  ccensor = list(
      type = "administrative_plus_random", # none | administrative | random | administrative_plus_random
      administrative_time = 90,
      random = list(dist = "exponential", rate = 0.0018)
      # future: target_censor, calibrate, etc.
    ),
  seed = NULL
  
)

#tenSeeds <- .Random.seed[1:10]
basic_RandomCensoring_sim <- list()

for(i in 1:10){
basic_RandomCensoring_sim[[i]] <- sim_generate(basic_RandomCensoring,
                         custom_design_matrix = scale(example_dm,
                                                      center = TRUE,
                                                      scale = FALSE),
                         custom_predictor_set = list(scenario_default_sim$truth$cause1,
                                                  scenario_default_sim$truth$cause2),
                         custom_betas = 
                list(scenario_default_sim$truth$effects$effects1$betas1,
                     scenario_default_sim$truth$effects$effects2$betas2),
                         seed = tenSeeds[i])
}
```

```{r}
outcome_simulation_diagnostic_plots(basic_RandomCensoring_sim[[2]], shapes = c(1,1), scales = c(0.0139,0.0139))
```

```{r}
summary_tables <- list()
for(i in 1:10){
temp1<-outcome_simulation_diagnostic_plots(basic_RandomCensoring_sim[[i]], shapes = c(1,1), scales = c(0.0139,0.0139))
temp2<- temp1$ObservedOutcomeSummary
temp2$SimNumber = rep(i, nrow(temp2))
summary_tables[[i]]<- temp2
rm(temp2)
}
basic_RandomCensoring_Summary = do.call("rbind", summary_tables)

basic_RandomCensoring_Summary
```

```{r}
basic_RandomCensoring_Summary %>%
  group_by(status)%>%
  summarise(Mean_Proportion = mean(Proportion, na.rm = T),
            SD_Proportion = sd(Proportion, na.rm = T))
```

```{r}
basic_RandomCensoring_Summary %>%
  select(status, Observed_TTE_median, Latent_TTE_median)%>%
  tidyr::pivot_longer(cols = c(Observed_TTE_median, Latent_TTE_median),
                      names_to = "True",
                      values_to = "TTE_Median")%>%
  ungroup()%>%
  ggplot(aes(x = status, y = TTE_Median, color = True))+
  geom_boxplot()+
  ylab("Median Time to Event")+
  xlab("Outcome")+
  scale_color_manual(name = "",
                     values = c("red", "blue"),
                     labels = c("Latent", "Observed"))+
  theme_classic()+
  ggtitle("Observed and Latent Event Times Across 10 Runs")
```

# Advanced scenarios

## Adv, no censor

```{r}
adv_NoCensoring<- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
   truth = list(#share_active_set = 2,
                shared_multiplier = c(0.5, 0.2)#,
                #sparsity = 0.05,
                #beta = list(dist = "normal", mean = 0, sd = c(0.3, 0.3)),
                #target_eta_sd = c(0.5, 0.5)), # variable selection
   ),
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 0.8, scale = -log(0.5)/(50^0.8)),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1.1, scale = -log(0.5)/(50^1.1)),
        link = "PH"
      )), # risk 
  censor = list(type = "none",
                administrative_time = 90), #type of censoring
  seed = NULL
  
)

tenSeeds <- .Random.seed[1:10]
adv_NoCensoring_sim <- list()

for(i in 1:10){
adv_NoCensoring_sim[[i]] <- sim_generate(adv_NoCensoring,
                         custom_design_matrix = scale(example_dm,
                                                      center = TRUE,
                                                      scale = FALSE),
                         custom_predictor_set = list(scenario_default_sim$truth$cause1,
                                                  scenario_default_sim$truth$cause2),
                         custom_betas = 
                list(scenario_default_sim$truth$effects$effects1$betas1,
                     scenario_default_sim$truth$effects$effects2$betas2),
                         seed = tenSeeds[i])
}



```

```{r}
outcome_simulation_diagnostic_plots(adv_NoCensoring_sim[[2]], shapes = c(0.8,1.1), scales = c(0.03031443,0.009374723))
```

```{r}

summary_tables <- list()
for(i in 1:10){
temp1<-outcome_simulation_diagnostic_plots(adv_NoCensoring_sim[[i]], shapes = c(0.8,1.1), scales = c( 0.03031443,0.009374723))
temp2<- temp1$ObservedOutcomeSummary
temp2$SimNumber = rep(i, nrow(temp2))
summary_tables[[i]]<- temp2
rm(temp2)
}
adv_NoCensoring_Summary = do.call("rbind", summary_tables)

adv_NoCensoring_Summary

adv_NoCensoring_Summary %>%
  group_by(status)%>%
  summarise(Mean_Proportion = mean(Proportion, na.rm = T),
            SD_Proportion = sd(Proportion, na.rm = T))

adv_NoCensoring_Summary %>%
  select(status, Observed_TTE_median, Latent_TTE_median)%>%
  tidyr::pivot_longer(cols = c(Observed_TTE_median, Latent_TTE_median),
                      names_to = "True",
                      values_to = "TTE_Median")%>%
  ungroup()%>%
  filter(status != "Censored")%>%
  ggplot(aes(x = status, y = TTE_Median, color = True))+
  geom_boxplot()+
  ylab("Median Time to Event")+
  xlab("Outcome")+
  scale_color_manual(name = "",
                     values = c("red", "blue"),
                     labels = c("Latent", "Observed"))+
  theme_classic()+
  ggtitle("Observed and Latent Event Times Across 10 Runs")
```

## Adv, admin censor

```{r}
adv_AdminCensoring<- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
   truth = list(#share_active_set = 2,
                shared_multiplier = c(0.5, 0.2)#,
                #sparsity = 0.05,
                #beta = list(dist = "normal", mean = 0, sd = c(0.3, 0.3)),
                #target_eta_sd = c(0.5, 0.5)), # variable selection
   ),
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 0.8, scale = -log(0.5)/(50^0.8)),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1.1, scale = -log(0.5)/(50^1.1)),
        link = "PH"
      )), # risk 
  censor = list(type = "administrative",
                administrative_time = 90), #type of censoring
  seed = NULL
  
)

adv_AdminCensoring_sim <- list()

for(i in 1:10){
adv_AdminCensoring_sim[[i]] <- sim_generate(adv_AdminCensoring,
                         custom_design_matrix = scale(example_dm,
                                                      center = TRUE,
                                                      scale = FALSE),
                         custom_predictor_set = list(scenario_default_sim$truth$cause1,
                                                  scenario_default_sim$truth$cause2),
                         custom_betas = 
                list(scenario_default_sim$truth$effects$effects1$betas1,
                     scenario_default_sim$truth$effects$effects2$betas2),
                         seed = tenSeeds[i])
}



summary_tables <- list()
for(i in 1:10){
temp1<-outcome_simulation_diagnostic_plots(adv_AdminCensoring_sim[[i]], shapes = c(0.8,1.1), scales = c( 0.03031443,0.009374723))
temp2<- temp1$ObservedOutcomeSummary
temp2$SimNumber = rep(i, nrow(temp2))
summary_tables[[i]]<- temp2
rm(temp2)
}
adv_AdminCensoring_Summary = do.call("rbind", summary_tables)

adv_AdminCensoring_Summary

adv_AdminCensoring_Summary %>%
  group_by(status)%>%
  summarise(Mean_Proportion = mean(Proportion, na.rm = T),
            SD_Proportion = sd(Proportion, na.rm = T))

adv_AdminCensoring_Summary %>%
  select(status, Observed_TTE_median, Latent_TTE_median)%>%
  tidyr::pivot_longer(cols = c(Observed_TTE_median, Latent_TTE_median),
                      names_to = "True",
                      values_to = "TTE_Median")%>%
  ungroup()%>%
  ggplot(aes(x = status, y = TTE_Median, color = True))+
  geom_boxplot()+
  ylab("Median Time to Event")+
  xlab("Outcome")+
  scale_color_manual(name = "",
                     values = c("red", "blue"),
                     labels = c("Latent", "Observed"))+
  theme_classic()+
  ggtitle("Observed and Latent Event Times Across 10 Runs")
```

## Adv, random censor

```{r}
adv_RandomCensoring<- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
   truth = list(#share_active_set = 2,
                shared_multiplier = c(0.5, 0.2)#,
                #sparsity = 0.05,
                #beta = list(dist = "normal", mean = 0, sd = c(0.3, 0.3)),
                #target_eta_sd = c(0.5, 0.5)), # variable selection
   ),
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 0.8, scale = -log(0.5)/(50^0.8)),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1.1, scale = -log(0.5)/(50^1.1)),
        link = "PH"
      )), # risk 
  censor = list(
      type = "administrative_plus_random", # none | administrative | random | administrative_plus_random
      administrative_time = 90,
      random = list(dist = "exponential", rate = 0.0018)
      # future: target_censor, calibrate, etc.
    ),
  seed = NULL
  
)

adv_RandomCensoring_sim <- list()

for(i in 1:10){
adv_RandomCensoring_sim[[i]] <- sim_generate(adv_RandomCensoring,
                         custom_design_matrix = scale(example_dm,
                                                      center = TRUE,
                                                      scale = FALSE),
                         custom_predictor_set = list(scenario_default_sim$truth$cause1,
                                                  scenario_default_sim$truth$cause2),
                         custom_betas = 
                list(scenario_default_sim$truth$effects$effects1$betas1,
                     scenario_default_sim$truth$effects$effects2$betas2),
                         seed = tenSeeds[i])
}



summary_tables <- list()
for(i in 1:10){
temp1<-outcome_simulation_diagnostic_plots(adv_RandomCensoring_sim[[i]], shapes = c(0.8,1.1), scales = c( 0.03031443,0.009374723))
temp2<- temp1$ObservedOutcomeSummary
temp2$SimNumber = rep(i, nrow(temp2))
summary_tables[[i]]<- temp2
rm(temp2)
}
adv_RandomCensoring_Summary = do.call("rbind", summary_tables)

adv_RandomCensoring_Summary

adv_RandomCensoring_Summary %>%
  group_by(status)%>%
  summarise(Mean_Proportion = mean(Proportion, na.rm = T),
            SD_Proportion = sd(Proportion, na.rm = T))

adv_RandomCensoring_Summary %>%
  select(status, Observed_TTE_median, Latent_TTE_median)%>%
  tidyr::pivot_longer(cols = c(Observed_TTE_median, Latent_TTE_median),
                      names_to = "True",
                      values_to = "TTE_Median")%>%
  ungroup()%>%
  #filter(status != "Censored")%>%
  ggplot(aes(x = status, y = TTE_Median, color = True))+
  geom_boxplot()+
  ylab("Median Time to Event")+
  xlab("Outcome")+
  scale_color_manual(name = "",
                     values = c("red", "blue"),
                     labels = c("Latent", "Observed"))+
  theme_classic()+
  ggtitle("Observed and Latent Event Times Across 10 Runs")
```

# Modeling

```{r}
library(fastcmprsk, lib.loc = "/Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library")
library(BhGLM, lib.loc = "/Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library")
```

```{r}
cs_oracle_func <- function(df){
    
    cox_true <- coxph(
    Surv(timing, status) ~ .,
    data = df
  )

  cox_true_sum <- summary(cox_true)

  cs_mod_res <- cox_true_sum$coefficients %>%
    as.data.frame()%>%
    rownames_to_column("Variable")%>%
    filter(`Pr(>|z|)` < 0.05)%>%
    dplyr::select(Variable, coef)

  names(cs_mod_res)<- c("Variable", "Estimate")

  cs_mod_res
  }
```

```{r}
cs_ssl_func <- function(x,y){
  # 2. Define the grid for the scale parameters (s0 and s1)
s0_grid <- seq(0.005, 0.1, length.out = 20)    # Spike scales (small values)
s1_grid <- c(0.5)      # Slab scales (larger values)

# Pre-allocate variables to store results
results <- data.frame()
best_cindex <- -Inf
best_ss <- c(NA, NA)

# 3. Create a fixed fold ID for consistent cross-validation splits
nobs <- nrow(x)
nfolds <- 10
#set.seed(123) # For reproducibility
foldid <- sample(rep(1:nfolds, length.out = nobs))

# 4. Run the grid search
for (s0 in s0_grid) {
  for (s1 in s1_grid) {
    
    # The spike scale must be smaller than the slab scale
    if (s0 >= s1) next 
    
    # Fit the bmlasso model using the current scale values
    #fit <- bmlasso(x = x, y = y, family = "cox", ss = c(s0, s1), verbose = FALSE)
    fit <- BhGLM::bmlasso(x,
                     y,
                     family = "cox",
                     ss = c(s0, s1))
    
    # Run cross-validation using the fixed folds
    cv_res <- cv.bh(fit, foldid = foldid, verbose = FALSE)
    
    # Extract performance metrics for Cox models (Cindex and deviance)
    cindex <- cv_res$measures["Cindex"]
    dev <- cv_res$measures["deviance"]
    
    # Store the results of this iteration
    results <- rbind(results, data.frame(s0 = s0, s1 = s1, 
                                         Cindex = cindex, 
                                         Deviance = dev))
    
    # Check if this is the best model so far (maximizing Concordance Index)
    if (cindex > best_cindex) {
      best_cindex <- cindex
      best_ss <- c(s0, s1)
    }
  }
}


# 6. Fit the final model using the optimized scale values
final_model <- bmlasso(x,
                     y, family = "cox", ss = best_ss)

final_model
} 
```

```{r}
psdh_ssl_func <- function(x,y){
  fit <- fit_ssl_psdh(x,y)
  
  tunes <- tune_ssl_psdh(fit, 
                         seq(0.01, 0.7, 0.05),
                         seq(0.4, 0.9, 0.05)
                         )
  max_tune <- tunes %>%
    filter(score_mean == max(tunes$score_mean,na.rm = T))
  
  final_fit <- fit_ssl_psdh(x,y, ss = c(max_tune$s0, max_tune$s1))
  
  final_fit
  
}
```

```{r}
model_sim_func <- function(sim){
  # Predictors
  df <- as.data.frame(sim$X)
  event1_df <- as.data.frame(sim$X[,sim$truth$cause1])
  event2_df <- as.data.frame(sim$X[,sim$truth$cause2])

  # Cause specific hazard with true variables
  cs_event1_df <- event1_df
  cs_event1_df$timing <- sim$time
  cs_event1_df$status <- sim$status
  cs_event1_df$status[cs_event1$status == 2] <- 0
  
  cs_event2_df <- event2_df
  cs_event2_df$timing <- sim$time
  cs_event2_df$status <- sim$status
  cs_event2_df$status[cs_event2$status == 1] <- 0
  cs_event2_df$status[cs_event2$status == 2] <- 1
  
  # PSDH 
  psdh_timing <- sim$time
  psdh_status1 <- sim$status
  psdh_status2 <- sim$status/2
  psdh_status2[psdh_timing2 == 0.5] <- 2
  
  
  
  cs_event1_oracle <- cs_oracle_func(cs_event1_df)
  cs_event2_oracle <- cs_oracle_func(cs_event2_df)
  
  # Competing risks hazard with true variables
  
  psdh_event1_oracle <- fastcmprsk::fastCrr(Crisk(sim$time, sim$status,
        cencode = 0, failcode = 1) ~ 
        as.matrix(event1_df),
        variance = T)
  
  psdh_event2_oracle <- fastcmprsk::fastCrr(Crisk(sim$time, sim$status,
        cencode = 0, failcode = 2) ~ 
        as.matrix(event2_df),
        variance = T)
  
  
  #Cause-specific 
  
  cs_event1_ssl <- cs_ssl_func(sim$X, Surv(cs_event1_df$timing, 
                                           cs_event1_df$status))
  
  cs_event2_ssl <- cs_ssl_func(sim$X, Surv(cs_event2_df$timing, 
                                           cs_event2_df$status))
  
  #PSDH
  psdh_event1_ssl <- psdh_ssl_func(sim$X, Surv(psdh_timing,
                                               psdh_status1))
  
  psdh_event2_ssl <- psdh_ssl_func(sim$X, Surv(psdh_timing,
                                               psdh_status2))
 
 # Wrapper
  # event1_res <- data.frame(Features = names(event1_df),
  #                              True_Effects = unname(death_effects$beta))%>%
  # dplyr::mutate(Overlap = ifelse(Features %in% selected_features$overlap, "Yes", ""))%>%
  # select(Features, Overlap, True_Effects)%>%
  # full_join(cs_oracle_death %>%
  #             dplyr::mutate(Features = Variable)%>%
  #             dplyr::mutate(CS_Oracle = Estimate)%>%
  #             dplyr::select(Features, CS_Oracle),
  #           by = "Features")%>%
  # full_join(data.frame(Features = names(death_effects$beta),
  #                      PSDH_Oracle = psdh_oracle_death$coef),
  #           by = "Features")%>%
  # full_join(data.frame(Features = rownames(cs_ssl_death$beta), 
  #                      CS_SSL = cs_ssl_death$beta[,1])%>%
  #             dplyr::filter(abs(CS_SSL) > 0),
  #           by = "Features")%>%
  # full_join(psdh_ssl_death$coefficients %>%
  #             dplyr::filter(abs(PSDH_SSL) > 0),
  #           by = "Features")
  # 
  # gvhd_res <- data.frame(Features = names(gvhd_effects$beta),
  #                              True_Effects = unname(gvhd_effects$beta))%>%
  # dplyr::mutate(Overlap = ifelse(Features %in% selected_features$overlap, "Yes", ""))%>%
  # select(Features, Overlap, True_Effects)%>%
  # full_join(cs_oracle_gvhd %>%
  #             dplyr::mutate(Features = Variable)%>%
  #             dplyr::mutate(CS_Oracle = Estimate)%>%
  #             dplyr::select(Features, CS_Oracle),
  #           by = "Features")%>%
  # full_join(data.frame(Features = names(gvhd_effects$beta),
  #                      PSDH_Oracle = psdh_oracle_gvhd$coef),
  #           by = "Features")%>%
  # full_join(data.frame(Features = rownames(cs_ssl_gvhd$beta), 
  #                      CS_SSL = cs_ssl_gvhd$beta[,1])%>%
  #             dplyr::filter(abs(CS_SSL) > 0),
  #           by = "Features")%>%
  # full_join(psdh_ssl_gvhd$coefficients %>%
  #             dplyr::filter(abs(PSDH_SSL) > 0),
  #           by = "Features")
  # 
  # 
  # 
  # 
  # return(list("Death" = death_res,
  #             "GVHD" = gvhd_res))
  
  
  
  
}
```

```{r}
test <- as.data.frame(basic_AdminCensoring_sim[[1]]$X)
test$timing <- basic_AdminCensoring_sim[[1]]$time
test$status <- basic_AdminCensoring_sim[[1]]$status
```

```{r}

test_x <- as.matrix(basic_AdminCensoring_sim[[1]]$X)
test_y_time <- basic_AdminCensoring_sim[[1]]$time
test_y_status <- basic_AdminCensoring_sim[[1]]$status
test_y_status[test_y_status == 2] <- 0

f1 <- BhGLM::bmlasso(test_x,
                     Surv(test_y_time,test_y_status),
                     family = "cox",
                     ss = c(0.05, 0.5))

BhGLM::cv.bh(f1, nfolds = 10, ncv = 5)
```

```{r}
# 2. Define the grid for the scale parameters (s0 and s1)
s0_grid <- seq(0.005, 0.1, length.out = 20)    # Spike scales (small values)
s1_grid <- c(0.5)      # Slab scales (larger values)

# Pre-allocate variables to store results
results <- data.frame()
best_cindex <- -Inf
best_ss <- c(NA, NA)

# 3. Create a fixed fold ID for consistent cross-validation splits
nobs <- nrow(test_x)
nfolds <- 10
#set.seed(123) # For reproducibility
foldid <- sample(rep(1:nfolds, length.out = nobs))

# 4. Run the grid search
for (s0 in s0_grid) {
  for (s1 in s1_grid) {
    
    # The spike scale must be smaller than the slab scale
    if (s0 >= s1) next 
    
    # Fit the bmlasso model using the current scale values
    #fit <- bmlasso(x = x, y = y, family = "cox", ss = c(s0, s1), verbose = FALSE)
    fit <- BhGLM::bmlasso(test_x,
                     Surv(test_y_time,test_y_status),
                     family = "cox",
                     ss = c(s0, s1))
    
    # Run cross-validation using the fixed folds
    cv_res <- cv.bh(fit, foldid = foldid, verbose = FALSE)
    
    # Extract performance metrics for Cox models (Cindex and deviance)
    cindex <- cv_res$measures["Cindex"]
    dev <- cv_res$measures["deviance"]
    
    # Store the results of this iteration
    results <- rbind(results, data.frame(s0 = s0, s1 = s1, 
                                         Cindex = cindex, 
                                         Deviance = dev))
    
    # Check if this is the best model so far (maximizing Concordance Index)
    if (cindex > best_cindex) {
      best_cindex <- cindex
      best_ss <- c(s0, s1)
    }
  }
}

# 5. Review the optimal values
print("Cross-Validation Results:")
print(results)

cat("\nOptimal Scale Values (s0, s1):", best_ss, "\n")
cat("Best Cross-Validated C-Index:", best_cindex, "\n")

# 6. Fit the final model using the optimized scale values
final_model <- bmlasso(test_x,
                     Surv(test_y_time,test_y_status), family = "cox", ss = best_ss)
```
