---
title: "Visualizing the problem"
format: html
editor: visual
---

# Hypothesis #1

When we have an influential variable which affects both the event of interest and the competing risk, we believe that our model will be more likely to identify that influential variable than a model which ignores competing risks.

## Problem we are trying to solve:

A spike-and-slab cox model that treats competing risks as censored will have lower sensitivity if all of the variables which affect the event of interest also affect the competing risk.

## Showing the problem

We will create 2 scenarios.

1.  Scenario 1: The variables which affect the event of interest are completely separate from the variables which affect the competing event 

2.  Scenario 2: The variables which affect the event of interest are the same as the variables which affect the competing event 

## Simulating outcomes

We will keep the scenario simple by keeping the outcome distributions for the 2 events identical and no censoring.

### Scenario 1

First we will get defaults for 2 completely different sets of predictors.

```{r}
scenario1_default_setting<- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
   truth = list(share_active_set = 0,
                #shared_multiplier = c(0.5, 0.2)#,
                sparsity = 0.05,
                beta = list(dist = "normal", mean = 4, sd = c(1, 0.2)),
                target_eta_sd = c(1.2, 0.8) # variable selection
   ),
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      )), # risk 
  censor = list(type = "none",
                administrative_time = 65), #type of censoring
  seed = NULL
  
)

scenario1_default <- sim_generate(scenario1_default_setting,
                         custom_design_matrix = example_dm,
                         custom_predictor_set = NULL,
                         custom_betas = NULL,
                         seed = 9134)


```

```{r}
names(example_dm[,scenario1_default$truth$cause1])
scenario1_default$truth$effects$effects1$betas1
exp(scenario1_default$truth$effects$effects1$betas1)
```

```{r}
names(example_dm[,scenario1_default$truth$cause2])
scenario1_default$truth$effects$effects2$betas2
exp(scenario1_default$truth$effects$effects2$betas2)
```

```{r}
tenSeeds <- .Random.seed[1:100]
scenario1_sims <- list()

for(i in 1:100){
scenario1_sims[[i]] <- sim_generate(scenario1_default_setting,
                         custom_design_matrix = scale(example_dm,
                                                      center = TRUE,
                                                      scale = FALSE),
                         custom_predictor_set = list(scenario1_default$truth$cause1,
                                                  scenario1_default$truth$cause2),
                         custom_betas = 
                list(scenario1_default$truth$effects$effects1$betas1,
                     scenario1_default$truth$effects$effects2$betas2),
                         seed = tenSeeds[i])
}
```

### Scenario 2

Next we will get defaults for a single overlapping set of predictors. We'll generate the ratio of the effect on the event of interest to the competing risk by a uniform distribution between 0 and 1, ensuring that the effect is always greater on the competing risk than it is on the event of interest.

```{r}
scenario2_default_setting<- sim_spec (
  n = nrow(example_dm), # samples
  p = ncol(example_dm), # features
  #x = list(), # design matrix will be custom
   truth = list(share_active_set = 6,
                shared_multiplier = runif(6, 0, 1),
                sparsity = 0.05,
                beta = list(dist = "normal", mean = 4, sd = c(1, 0.2)),
                target_eta_sd = c(1.2, 0.8) # variable selection
   ),
  risks = list(
      cause1 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      ),
      cause2 = list(
        baseline = list(dist = "weibull", shape = 1, scale = 0.0139),
        link = "PH"
      )), # risk 
  censor = list(type = "none",
                administrative_time = 65), #type of censoring
  seed = NULL
  
)

scenario2_default <- sim_generate(scenario2_default_setting,
                         custom_design_matrix = example_dm,
                         custom_predictor_set = NULL,
                         custom_betas = NULL,
                         seed = 9134)
```

```{r}
names(example_dm[,scenario2_default$truth$cause1])
scenario2_default$truth$effects$effects1$betas1
exp(scenario2_default$truth$effects$effects1$betas1)
```

```{r}
names(example_dm[,scenario2_default$truth$cause2])
scenario2_default$truth$effects$effects2$betas2
exp(scenario2_default$truth$effects$effects2$betas2)
```

```{r}
scenario2_sims <- list()

for(i in 1:100){
scenario2_sims[[i]] <- sim_generate(scenario2_default_setting,
                         custom_design_matrix = scale(example_dm,
                                                      center = TRUE,
                                                      scale = FALSE),
                         custom_predictor_set = list(scenario2_default$truth$cause1,
                                                  scenario2_default$truth$cause2),
                         custom_betas = 
                list(scenario2_default$truth$effects$effects1$betas1,
                     scenario2_default$truth$effects$effects2$betas2),
                         seed = tenSeeds[i])
}
```

### Visualizing outcomes

We would expect to see approximately equal event rates across either scenario because the underlying shape and scale distributions are the same for both event types and there is no censoring.

```{r}
get_status_proportions <- function(sim_list) {
  # Loop through each simulation in the list
  results <- lapply(seq_along(sim_list), function(i) {
    # Extract the current iteration
    current_sim <- sim_list[[i]]
    
    # Calculate proportions (mean of a logical vector gives the proportion)
    # na.rm = TRUE is included just in case you have missing data
    p1 <- mean(current_sim$status == 1, na.rm = TRUE)
    p2 <- mean(current_sim$status == 2, na.rm = TRUE)
    
    # Return as a single-row data frame
    data.frame(
      iteration = i,
      prop_status_1 = p1,
      prop_status_2 = p2
    )
  })
  
  # Bind all the individual rows into one comprehensive data frame
  final_df <- do.call(rbind, results)
  return(final_df)
}
```

```{r}
scenario1_proportions <- get_status_proportions(scenario1_sims)

plot_scenario1_outcomes <- scenario1_proportions %>%
  # Make sure iteration is numeric (especially if you used purrr and it made it a character)
  mutate(iteration = as.numeric(iteration)) %>% 
  pivot_longer(
    cols = c(prop_status_1, prop_status_2), 
    names_to = "status_type", 
    values_to = "proportion"
  )

# 2. Create the plot
ggplot(plot_scenario1_outcomes, aes(x = iteration, y = proportion, color = status_type)) +
  geom_line(alpha = 0.7) +   # Add lines connecting the iterations
  geom_point(size = 1.5) +   # Add points for each specific iteration
  scale_y_continuous(limits = c(0, 1)) + # Proportions always fall between 0 and 1
  scale_color_manual(
    values = c("prop_status_1" = "#1b9e77", "prop_status_2" = "#d95f02"),
    labels = c("Status 1", "Status 2")
  ) +
  theme_minimal() +
  labs(
    title = "Stability of Event Rates Across Simulations: Scenario 1",
    subtitle = "Tracking Status 1 (Event of Interest) vs Status 2 (Competing Event) for each iteration",
    x = "Simulation Iteration",
    y = "Proportion of Events",
    color = "Event Status",
    caption = "100 iterations"
  ) +
  theme(legend.position = "bottom")
```

```{r}
scenario2_proportions <- get_status_proportions(scenario2_sims)

plot_scenario2_outcomes <- scenario2_proportions %>%
  # Make sure iteration is numeric (especially if you used purrr and it made it a character)
  mutate(iteration = as.numeric(iteration)) %>% 
  pivot_longer(
    cols = c(prop_status_1, prop_status_2), 
    names_to = "status_type", 
    values_to = "proportion"
  )

# 2. Create the plot
ggplot(plot_scenario2_outcomes, aes(x = iteration, y = proportion, color = status_type)) +
  geom_line(alpha = 0.7) +   # Add lines connecting the iterations
  geom_point(size = 1.5) +   # Add points for each specific iteration
  scale_y_continuous(limits = c(0, 1)) + # Proportions always fall between 0 and 1
  scale_color_manual(
    values = c("prop_status_1" = "#1b9e77", "prop_status_2" = "#d95f02"),
    labels = c("Status 1", "Status 2")
  ) +
  theme_minimal() +
  labs(
    title = "Stability of Event Rates Across Simulations: Scenario 2",
    subtitle = "Tracking Status 1 (Event of Interest) vs Status 2 (Competing Event) for each iteration",
    x = "Simulation Iteration",
    y = "Proportion of Events",
    color = "Event Status",
    caption = "100 iterations"
  ) +
  theme(legend.position = "bottom")
```

## Modeling

### Functions

```{r}
cs_ssl_sim_func <- function(sim){
  # Predictors
  df <- as.data.frame(sim$X)
  event1_df <- as.data.frame(sim$X[,sim$truth$cause1])
  event2_df <- as.data.frame(sim$X[,sim$truth$cause2])

  # Cause specific hazard with true variables
  cs_event1_df <- event1_df
  cs_event1_df$timing <- sim$time
  cs_event1_df$status <- sim$status
  cs_event1_df$status[cs_event1_df$status == 2] <- 0
  
  cs_event2_df <- event2_df
  cs_event2_df$timing <- sim$time
  cs_event2_df$status <- sim$status
  cs_event2_df$status[cs_event2_df$status == 1] <- 0
  cs_event2_df$status[cs_event2_df$status == 2] <- 1
  
  # # PSDH 
  # psdh_timing <- sim$time
  # psdh_status1 <- sim$status
  # psdh_status2 <- sim$status/2
  # psdh_status2[psdh_status2 == 0.5] <- 2
  
  
  
  # cs_event1_oracle <- cs_oracle_func(cs_event1_df)
  # cs_event2_oracle <- cs_oracle_func(cs_event2_df)
  
  
  
  # Competing risks hazard with true variables
  
  # psdh_event1_oracle <- psdh_oracle_func(event1_df, psdh_timing, psdh_status1, 1)
  # 
  # psdh_event2_oracle <- psdh_oracle_func(event2_df, psdh_timing, psdh_status2, 1)
  
  
  #Cause-specific 
  
  cs_event1_ssl <- cs_ssl_func(sim$X, Surv(cs_event1_df$timing, 
                                           cs_event1_df$status))
  
  cs_event2_ssl <- cs_ssl_func(sim$X, Surv(cs_event2_df$timing, 
                                           cs_event2_df$status))
  
  #PSDH
  # psdh_event1_ssl <- suppressWarnings(
  #   suppressMessages( psdh_ssl_func(sim$X, cbind(psdh_timing, 
  #                                          psdh_status1))
  #   ))
  # 
  # psdh_event2_ssl <- suppressWarnings(
  #   suppressMessages(psdh_ssl_func(sim$X, cbind(psdh_timing, 
  #                                          psdh_status2))
  #   ))
  

 
 # Wrapper
  
  models <- list(#"CS_Oracle_Event1" = cs_event1_oracle$CS_Oracle_Model,
                 #"CS_Oracle_Event2" = cs_event2_oracle$CS_Oracle_Model,
                 #"PSDH_Oracle_Event1" = psdh_event1_oracle$PSDH_Oracle_Model,
                 #"PSDH_Oracle_Event2" = psdh_event2_oracle$PSDH_Oracle_Model,
                 "CS_SSL_Event1" = cs_event1_ssl,
                 "CS_SSL_Event2" = cs_event2_ssl)#,
                 #"PSDH_SSL_Event1" = psdh_event1_ssl,
                 ##"PSDH_SSL_Event2" = psdh_event2_ssl)
  
  selected_features_summary_event1 <- data.frame(Features = names(event1_df),
                                True_Effects = sim$truth$effects$effects1$betas1)|>
    dplyr::mutate(Overlap = case_when(Features %in% names(df[,sim$truth$shared_predictors])~ "Yes", 
                                     TRUE ~ ""))|>
    select(Features, Overlap, True_Effects)|>
    # full_join(cs_event1_oracle$CS_Oracle_Summary ,
    #         by = "Features")|>
    # full_join(psdh_event1_oracle$PSDH_Oracle_Summary ,
    #         by = "Features")|>
    full_join(data.frame(Features = rownames(cs_event1_ssl$beta),
                        CS_SSL = cs_event1_ssl$beta[,1])|>
               dplyr::filter(abs(CS_SSL) > 0),
    by = "Features")#|>
    # full_join(data.frame(Features = colnames(df),
    #                     PSDH_SSL = psdh_event1_ssl$coef[,1])|>
    #            dplyr::filter(abs(PSDH_SSL) > 0),
    # by = "Features")
  
  selected_features_summary_event2 <- data.frame(Features = names(event2_df),
                                True_Effects = sim$truth$effects$effects2$betas2)|>
    dplyr::mutate(Overlap = case_when(Features %in% names(df[,sim$truth$shared_predictors])~ "Yes", 
                                     TRUE ~ ""))|>
    select(Features, Overlap, True_Effects)|>
    # full_join(cs_event2_oracle$CS_Oracle_Summary ,
    #         by = "Features")|>
    # full_join(psdh_event2_oracle$PSDH_Oracle_Summary ,
    #         by = "Features")|>
    full_join(data.frame(Features = rownames(cs_event2_ssl$beta),
                        CS_SSL = cs_event2_ssl$beta[,1])|>
               dplyr::filter(abs(CS_SSL) > 0),
    by = "Features")#|>
    # full_join(data.frame(Features = colnames(df),
    #                     PSDH_SSL = psdh_event2_ssl$coef[,1])|>
    #            dplyr::filter(abs(PSDH_SSL) > 0),
    # by = "Features")
  
  ret <- list("Models" = models,
              "Selected_Features_Event1" = selected_features_summary_event1,
              "Selected_Features_Event2" = selected_features_summary_event2
              )
  
  ret
  
 

  
}

parse_cs_ssl_simulation_results <- function(results_list, event) {
  str <- paste0("Selected_Features_Event", event)
  
  # Combine the list into one massive data frame, keeping track of the iteration
  all_data <- results_list %>%
  purrr::map(str)%>%
  bind_rows(.id = "iteration")
  
  # Define model names dynamically (excluding the metadata columns)
  model_names <- c("CS_SSL") #c("CS_Oracle",
                   #"PSDH_Oracle", 
                  # "CS_SSL", 
                  #"PSDH_SSL") 
  
  # Pivot the data so we have one row per model per feature per iteration
  long_data <- all_data %>%
     tidyr::pivot_longer(cols = all_of(model_names), names_to = "model", values_to = "estimate")%>%
     mutate(
       # 1. Logic for Signal Detection
       is_true_effect = !is.na(True_Effects),
       is_detected    = !is.na(estimate),
       is_special     = !is.na(Overlap) & Overlap == "Yes",
       
       # 2. Logic for Sign Consistency
       # Only relevant if both true_effect and estimate exist
       correct_sign = ifelse(is_true_effect & is_detected, 
                             sign(True_Effects) == sign(estimate), 
                             NA)
     )%>%
    transform(estimate = ifelse(is.na(estimate), 0, estimate))%>%
    transform(True_Effects = ifelse(is.na(True_Effects), 0, True_Effects))
  
  #print(head(long_data))
  
#   # Calculate Metrics per Model per Iteration
  summary_stats <- long_data %>%
    group_by(iteration, model) %>%
    summarise(
      sensitivity = sum(is_detected & is_true_effect) / sum(is_true_effect),
      specificity = sum(!is_detected & !is_true_effect) / sum(!is_true_effect),

      # Precision/Detection for 'Special' features
      special_recall = sum(is_detected & is_special) / sum(is_special),

      # Sign Accuracy (among true signals that were actually detected)
      sign_accuracy = sum(correct_sign, na.rm = TRUE) / sum(is_detected & is_true_effect),
      
      MSE = mean((estimate - True_Effects)^2),

      .groups = 'drop'
    )
  
  #print(head(summary_stats))

  return(summary_stats)
 }
```

### Running Simulation

```{r}
library(future.apply)
```

```{r}
# 1. Set up the parallel backend
# 'multisession' works on all operating systems and creates background R sessions.
# We leave one core free for system stability.
plan(multisession, workers = availableCores() - 1) 
```

```{r}

# ==========================================
# Scenario 1
# ==========================================

# Replace the for-loop with future_lapply
scenario1_output <- future_lapply(1:100, function(i) {
  # Note: Standard print() won't show up in the main console during parallel execution
  #message(paste("Starting Scenario 1, iteration", i, "at", Sys.time()))
  
  # Run the simulation
  cs_ssl_sim_func(scenario1_sims[[i]])
  
}, future.seed = TRUE) # IMPORTANT: Use future.seed = TRUE for safe random number generation in simulations

# Keep your existing parsing code...
scenario1a_results <- parse_cs_ssl_simulation_results(scenario1_output, "1")
scenario1a_results$Event = rep("Event1", nrow(scenario1a_results))
scenario1b_results <- parse_cs_ssl_simulation_results(scenario1_output, "2")
scenario1b_results$Event = rep("Event2", nrow(scenario1b_results))
scenario1_results <- rbind(scenario1a_results, scenario1b_results)
scenario1_results$Scenario = "Disjoint Predictors"


# ==========================================
# Scenario 2
# ==========================================

# Replace the for-loop with future_lapply
scenario2_output <- future_lapply(1:100, function(i) {
  #message(paste("Starting Scenario 2, iteration", i, "at", Sys.time()))
  
  # Run the simulation
  cs_ssl_sim_func(scenario2_sims[[i]])
  
}, future.seed = TRUE)

# Keep your existing parsing code...
scenario2a_results <- parse_cs_ssl_simulation_results(scenario2_output, "1")
scenario2a_results$Event = rep("Event1", nrow(scenario2a_results))
scenario2b_results <- parse_cs_ssl_simulation_results(scenario2_output, "2")
scenario2b_results$Event = rep("Event2", nrow(scenario2b_results))
scenario2_results <- rbind(scenario2a_results, scenario2b_results)
scenario2_results$Scenario = "Mutual Predictors"


# ==========================================
# Joining
# ==========================================
hypothesis1 <- rbind(scenario1_results, scenario2_results)


# plan(sequential)
```

## Plotting

```{r}
ggplot(hypothesis1 %>%
         filter(Event == "Event1"), aes(x = sensitivity,
                        color = Scenario,
                        fill = Scenario))+
  geom_density(alpha = 0.2)+
  labs(title = "Distribution of sensitivity scores across simulations for \n
detecting prognostic variables for the event of interest: \n
cause-specific spike-and-slab LASSO cox model",
       x = "Sensitivity score",
       y = "Proportion of simulations",
       caption = "100 iterations")+
  theme_classic()
```

```{r}
ggplot(hypothesis1 %>%
         filter(Event == "Event1"), aes(x = sensitivity,
                        color = Scenario,
                        fill = Scenario))+
  geom_histogram(alpha = 0.2, position = "dodge")+
  labs(title = "Distribution of sensitivity scores across simulations for \n
detecting prognostic variables for the event of interest: \n
cause-specific spike-and-slab LASSO cox model",
       x = "Sensitivity score",
       y = "Number of simulations",
       caption = "100 iterations")+
  theme_classic()
```

```{r}
ggplot(hypothesis1 %>%
         filter(Event == "Event2"), aes(x = sensitivity,
                        color = Scenario,
                        fill = Scenario))+
  geom_density(alpha = 0.2)+
  labs(title = "Distribution of sensitivity scores across simulations for \n
detecting prognostic variables for the competing event: \n
cause-specific spike-and-slab LASSO cox model",
       x = "Sensitivity score",
       y = "Proportion of simulations",
       caption = "100 iterations")+
  theme_classic()
```

```{r}
ggplot(hypothesis1 %>%
         filter(Event == "Event2"), aes(x = sensitivity,
                        color = Scenario,
                        fill = Scenario))+
  geom_histogram(alpha = 0.2, position = "dodge")+
  labs(title = "Distribution of sensitivity scores across simulations for \n
detecting prognostic variables for the competing event: \n
cause-specific spike-and-slab LASSO cox model",
       x = "Sensitivity score",
       y = "Number of simulations",
       caption = "100 iterations")+
  theme_classic()
```
