---
title: "Solution Paths"
format: html
editor: visual
---

# Loading functions

```{r}
library(tidyverse)
library(fastcmprsk, lib.loc = "/Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library")
library(survival)

# source("/Users/sophiehuebler/Documents/bhCRR/R/cv_ssl_psdh.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/fit_ssl_psdh.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/update_betas.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/expected_inclusion_probs.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/expected_penalty_weights.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/update_mixture_prob.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/helpers.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/tune_ssl_psdh.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/measure_ssl_psdh.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/predict_from_ssl_psdh.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/generate_foldid.r")
# source("/Users/sophiehuebler/Documents/bhCRR/R/dedupe_warnings.r")
# source("/Users/sophiehuebler/Documents/bhCRR/src/cv_fastcrrp.cpp")
# source("/Users/sophiehuebler/Documents/bhCRR/R/cv_fastCrrp_cpp.r")
Rcpp::compileAttributes("~/Documents/bhCRR")
devtools::load_all("~/Documents/bhCRR", recompile = TRUE)
```

# Generating data

```{r}
###########Generate Data Functions ##########################################

easy_sim <- function(nobs =50,
                     beta1,
                     beta2){
  sim <- simulateTwoCauseFineGrayModel(nobs = 50,
                                       beta1 = beta1,
                                       beta2 = beta2,
                                       X = NULL,
                                       u.min = 100,
                                       u.max = 100,
                                       p = 0.5,
                                       returnX = TRUE)

  sim_data <- cbind(data.frame(ID = 1:nobs,
                               TTE = sim$ftime,
                               Status = as.integer(sim$fstatus)),
                    sim$X)

  names(sim_data)<- c("ID", "TTE", "Status", paste0("X_", 1:(ncol(sim_data)-3)))

  x <- as.matrix(sim_data %>%select(starts_with("X")))
  y <- as.matrix(sim_data %>%
                   mutate(Status = as.numeric(Status))%>%
                   select(TTE, Status))

  return(list("dataframe" = sim_data,
              "x" = x,
              "y" = y))
}

easy_sim_wrapper <- function(x,
                             nobs = 50,
                             beta1_active = c(0.40, - 0.50, 0.60, 0.75, - 0.80), 
                             beta2_active = c(0, 0.3, 0, 0, -0.2)){
data <- easy_sim(nobs,
                     c(beta1_active, rep(0,x-length(beta1_active))),
                     c(beta2_active, rep(0,x-length(beta1_active)))
)


fg_data <- data$dataframe
fg_data$Status_factor <- factor(fg_data$Status,
                           levels = c(0, 1, 2),
                           labels = c("censor", "event", "competing"))
fg_data_format <- finegray(Surv(TTE, Status_factor) ~ ., data = fg_data, etype = "event")

sparse_fg_formula <- paste("Surv(fgstart, fgstop, fgstatus) ~ ",
                           paste0("X_", 1:length(beta1_active), collapse = " + "))

full_fg_formula <- paste("Surv(fgstart, fgstop, fgstatus) ~ ",
                           paste0("X_", 1:x, collapse = " + "))

sparse_fg_model <- coxph(as.formula(sparse_fg_formula),
                  weight = fgwt,
                  data = fg_data_format)

if(x < nobs){
  full_fg_model <- coxph(as.formula(full_fg_formula),
                  weight = fgwt,
                  data = fg_data_format)

  full_fg_ests <- as.vector(full_fg_model$coefficients)
}else{
  full_fg_ests <- rep(NA, x)
}





comparisons <- data.frame(Name = c(paste("V", 1:x, sep = "")),
                   True_beta = c(beta1_active,
                                 rep(0, x-length(beta1_active))),
                   Sparse_FG_beta = c(as.vector(sparse_fg_model$coefficients),
                                 rep(0, x-length(beta1_active))),
                   Full_FG_beta = full_fg_ests
)

return(list(data = data,
            comparisons=comparisons))
}
```

## Basic Scenarios

```{r}
basic5 <- easy_sim_wrapper(5, nobs = 100) 
basic25 <- easy_sim_wrapper(25, nobs = 100)
basic100 <- easy_sim_wrapper(101, nobs = 100)

basic5_large <- easy_sim_wrapper(5, nobs = 100,
                                 beta1_active = c(2.40, - 2.50, 2.60, 2.75, - 2.80), 
                                 beta2_active = c(0, 1.3, 0, 0, -2.2)) 
basic25_large <- easy_sim_wrapper(25, nobs = 100,
                                 beta1_active = c(2.40, - 2.50, 2.60, 2.75, - 2.80), 
                                 beta2_active = c(0, 1.3, 0, 0, -2.2)) 
basic100_large <- easy_sim_wrapper(101, nobs = 100,
                                 beta1_active = c(2.40, - 2.50, 2.60, 2.75, - 2.80), 
                                 beta2_active = c(0, 1.3, 0, 0, -2.2)) 
```

## Visualizing single models

```{r}
mod_pips_func <- function(ss, scenario, supply_mod = NULL){
  
  if(is.null(supply_mod)){
  mod <- fit_ssl_psdh(scenario$data$x,scenario$data$y, ss = ss,
                            initial_sparsity = 0.5)
  }else{
    mod <- supply_mod
  }
  
pips <- as.data.frame(mod$pips_path)
names(pips) <- NULL
names(pips) <- c(paste0("Iteration_", 0:(ncol(pips)-1)))
pips$Beta <- paste0("V_", 1:nrow(scenario$comparisons))
pips <- pips %>%
  pivot_longer(cols = -Beta, values_to = "Est", names_to = "Iteration",
               names_prefix = "Iteration_", 
               names_transform = list(Iteration = as.numeric))

p1 <- ggplot(pips%>% 
         filter(!(Beta %in% c("V_1", "V_2", "V_3", "V_4", "V_5" ))),
        aes(x = Iteration, y = Est, color = Beta))+
  geom_point()+
  geom_line(aes(group = Beta))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  ggtitle("Inclusion Probabilities: Inactive")

print(p1)

p2 <- ggplot(pips %>% 
         filter(Beta %in% c("V_1", "V_2", "V_3", "V_4", "V_5" )),
          aes(x = Iteration, y = Est, color = Beta))+
  geom_point()+
  geom_line(aes(group = Beta))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  ggtitle("Inclusion Probabilities: Active")

print(p2)

betas <- as.data.frame(mod$betas_path)
names(betas) <- NULL
names(betas) <- c(paste0("Iteration_", 0:(ncol(betas)-1)))
betas$Beta <- paste0("V_", 1:nrow(scenario$comparisons))
betas <- betas %>%
  pivot_longer(cols = -Beta, values_to = "Est", names_to = "Iteration",
               names_prefix = "Iteration_", 
               names_transform = list(Iteration = as.numeric))

betas <- betas %>%
  left_join(scenario$comparisons %>% 
                             mutate(Beta = str_replace(Name, "^(V)(\\d+)$", "\\1_\\2")), 
                            by = c("Beta"))

p3 <- ggplot(betas%>% 
         filter(!(Beta %in% c("V_1", "V_2", "V_3", "V_4", "V_5" ))),
        aes(x = Iteration, y = Est, color = Beta))+
  geom_point()+
  geom_line(aes(group = Beta))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  ggtitle("Betas: Inactive")

print(p3)


p4 <- ggplot(betas %>% 
         filter(Beta %in% c("V_1", "V_2", "V_3", "V_4", "V_5" )),
          aes(x = Iteration, y = Est, color = Beta))+
  geom_point()+
  geom_line(aes(group = Beta))+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  ggtitle("Betas: Active")+
  geom_hline(aes(yintercept = True_beta, color = Beta), linetype = "dashed", alpha = 0.5)+
  geom_hline(aes(yintercept = Sparse_FG_beta, color = Beta), alpha = 0.5)+
  labs(caption = "True estimate in dashed line, sparse fine and gray estimate in solid line")

print(p4)

invisible(mod)


}
```

## Reasonable base models

```{r}
mod_basic5 <- mod_pips_func(ss = c(0.005, 1000), basic5, supply_mod = NULL)
```

```{r}
mod_basic25 <- mod_pips_func(ss = c(0.03, 1000), basic25, supply_mod = NULL)
```

```{r}
mod_basic25_large <- mod_pips_func(ss = c(0.03, 1000), basic25_large, supply_mod = NULL)
```

```{r}
mod_basic100 <- mod_pips_func(ss = c(0.03, 1000), basic100)
```

```{r}
mod_basic100 <- mod_pips_func(ss = c(0.03, 1000), basic100_large)
```

# Solution path

```{r}

solution_path_func <- function(scale0, scale1, scenario,
                               cindex = FALSE,
                               eval_time = NULL){
 # fit_ssl_psdh is called exactly ONCE per (s0, s1) combo; the resulting
 # object is reused for both the coefficient path and (optionally) the
 # Wolbers IPCW C-index.
 temp <-  fit_ssl_psdh(scenario$data$x,scenario$data$y, ss = c(scale0, scale1),
                            initial_sparsity = 0.5)

 ret <- as.vector(temp$final_model_object$coef)
 names(ret) <- paste0("V", seq_along(ret))

 if(cindex){
   # Default evaluation horizon: latest observed time in the scenario.
   et <- if(is.null(eval_time)) max(scenario$data$y[, 1]) else eval_time
   risk <- predict_from_ssl_psdh(temp,
                                 newx = scenario$data$x,
                                 prediction_time = et)
   ci <- measure_ssl_psdh(y_true = scenario$data$y,
                          risk_score = risk,
                          evaluation_time = et)
   ret <- c(ret, wolbers_cindex = ci)
 }

 ret

}


solution_path_wrapper_s0 <- function(scale1,
                                     path = NULL,
                                  start = NULL, end = NULL, granularity = NULL,
                                  scenario,
                                  cindex = FALSE,
                                  eval_time = NULL){
  
if(is.null(path)){
  path <- seq(start, end, granularity)
}else{
  path<-path[path<scale1]
}
  suppressWarnings({
    res_1 <- sapply(path, function(x) solution_path_func(x, scale1, scenario,
                                                         cindex = cindex,
                                                         eval_time = eval_time))
    })

res <- as.data.frame(t(res_1)) %>%
  mutate(s1 = scale1)
res$s0 = path

# Split the C-index off before pivoting the coefficient columns so it stays
# one value per (s0, s1) combo.
if(cindex){
  cindex_df <- res %>% select(s0, s1, wolbers_cindex)
  res <- res %>% select(-wolbers_cindex)
}

res <- res %>%
  pivot_longer(cols = starts_with("V"),
               names_to = "Beta", values_to = "Estimate")

res <- res %>%
  left_join(scenario$comparisons, by = c("Beta" = "Name"))%>%
  select(s0, s1, Beta, True_beta, Sparse_FG_beta, Full_FG_beta, Estimate)

if(all(is.na(res$Full_FG_beta))){
  res <- res %>% select(-Full_FG_beta)
}

if(cindex){
  res <- res %>% left_join(cindex_df, by = c("s0", "s1"))
}

res

}

solution_path_wrapper <- function(scale1_seq,
                                  scale0_path = NULL,
                                  scale0_start = NULL, scale0_granularity = NULL,
                                  scenario,
                                  cindex = FALSE,
                                  eval_time = NULL){

 dfs <- lapply(scale1_seq, function(x) solution_path_wrapper_s0(x,
                                     path = scale0_path,
                                  start = scale0_start, 
                                  end = x, granularity = scale0_granularity,
                                  scenario,
                                  cindex = cindex,
                                  eval_time = eval_time))

 ret <- do.call("rbind", dfs)

 ret
}
```

# Basic 5

```{r}
basic5_grid <- solution_path_wrapper(scale1_seq = c(1, 100, 1000),
                                     #scale0_path = seq(0.001, 0.1, .005),
                                     scale0_path = 1/(seq(0.0001, 100, 10)),
                                     scale0_start = 0.001,
                                     scale0_granularity = 0.05,
                                     scenario = basic5, 
                                     cindex = TRUE)


```

```{r}
ggplot(basic5_grid, 
       aes(x = as.factor(s0), y = as.factor(s1), fill = wolbers_cindex)) +
  geom_tile(color = "white") +
  # viridis is great for continuous gradients and colorblind accessibility
  scale_fill_viridis_c(option = "plasma") + 
  theme_minimal() +
  labs(
    title = "Hyperparameter Landscape",
    #subtitle = "Look for broad regions of high scores, not just isolated peaks",
    x = "Parameter s0",
    y = "Parameter s1",
    fill = "C-index"
  )+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

```{r}
ggplot(basic5_grid %>%
         filter(Beta %in% c("V1", "V2", "V3", "V4", "V5" ))%>%
         filter(s1 == 1000)
       )+
  geom_point(aes(x=1/s0, y = Estimate, color = Beta))+
  geom_line(aes(x=1/s0, y = Estimate, color = Beta), alpha = 0.5)+
  geom_hline(aes(yintercept = True_beta, color = Beta), linetype = "dashed")+
  geom_hline(aes(yintercept = Sparse_FG_beta, color = Beta), alpha = 0.5)+
  labs(caption = "True estimate in dashed line, sparse fine and gray estimate in solid line")
```

## Observations

-   The CV index is max when s0 scale is minimal. This makes sense, as it correlates to little shrinkage. There should be no shrinkage in this scenario.

-   The estimates trend towards the fine and gray estimates (solid lines) rather than the true values (dashed lines).

-   There is some shrinkage towards 0 even as s0 extends out in either direction. It is on the scale of 1/s1, so s1 should be about 1000 for it not to really affect our estimates.

# Basic 25

```{r}
basic25_grid <- solution_path_wrapper(scale1_seq = c(100, 500, 1000),
                                     #scale0_path = seq(0.001, 0.1, .005),
                                     scale0_path = 1/(seq(0.0001, 100, 10)),
                                     scale0_start = 0.001,
                                     scale0_granularity = 0.05,
                                     scenario = basic25, 
                                     cindex = TRUE)


ggplot(basic25_grid %>%
         filter(Beta %in% c("V1", "V2", "V3", "V4", "V5" ))%>%
         filter(s1 == 1000)
       )+
  geom_point(aes(x=1/s0, y = Estimate, color = Beta))+
  geom_line(aes(x=1/s0, y = Estimate, color = Beta), alpha = 0.5)+
  geom_hline(aes(yintercept = True_beta, color = Beta), linetype = "dashed")+
  geom_hline(aes(yintercept = Sparse_FG_beta, color = Beta), alpha = 0.5)+
  labs(caption = "True estimate in dashed line, sparse fine and gray estimate in solid line",
      title = "Active Coefficients")

ggplot(basic25_grid %>%
         filter(!(Beta %in% c("V1", "V2", "V3", "V4", "V5" )))%>%
         filter(s1 == 1000)
       )+
  geom_point(aes(x=1/s0, y = Estimate, color = Beta))+
  geom_line(aes(x=1/s0, y = Estimate, color = Beta), alpha = 0.5)+
  geom_hline(aes(yintercept = True_beta, color = Beta), linetype = "dashed")+
  geom_hline(aes(yintercept = Sparse_FG_beta, color = Beta), alpha = 0.5)+
  labs(caption = "True estimate in dashed line, sparse fine and gray estimate in solid line",
      title = "Inactive Coefficients")
```

```{r}
target_betas <- c("V1", "V2", "V3", "V4", "V5")

ggplot() +
  # LAYER 1 (BOTTOM): Plot everything EXCEPT V1-V5 in light gray
  geom_line(
    data = subset(basic25_grid, !(Beta %in% target_betas) & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "darkgray"
  ) +
  geom_point(
    data = subset(basic25_grid, !(Beta %in% target_betas)& s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "darkgray"
  ) +
  
  # LAYER 2 (TOP): Plot ONLY V1-V5 in black so they draw over the gray
  geom_line(
    data = subset(basic25_grid, Beta %in% target_betas & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "black"
  ) +
  geom_point(
    data = subset(basic25_grid, Beta %in% target_betas & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "black"
  ) +
  
  # Set the x-axis ticks to match the exact s0 values
  scale_x_continuous(breaks = 1/unique(basic25_grid$s0)) +
  
  # Rotate the text 90 degrees and align it properly
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```

```{r}
basic25_grid_zoomed <- solution_path_wrapper(scale1_seq = c( 1000),
                                     #scale0_path = seq(0.001, 0.1, .005),
                                     scale0_path =seq(0.02, 0.04, .001),
                                     scale0_start = 0.001,
                                     scale0_granularity = 0.05,
                                     scenario = basic25, 
                                     cindex = TRUE)

ggplot(basic25_grid_zoomed, 
       aes(x = as.factor(s0), y = as.factor(s1), fill = wolbers_cindex)) +
  geom_tile(color = "white") +
  # viridis is great for continuous gradients and colorblind accessibility
  scale_fill_viridis_c(option = "plasma") + 
  theme_minimal() +
  labs(
    title = "Hyperparameter Landscape",
    #subtitle = "Look for broad regions of high scores, not just isolated peaks",
    x = "Parameter s0",
    y = "Parameter s1",
    fill = "C-index"
  )+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

# Basic 100

```{r}
basic100_grid <- solution_path_wrapper(scale1_seq = c( 1000),
                                     #scale0_path = seq(0.001, 0.1, .005),
                                     scale0_path = 1/(seq(0.0001, 100, 10)),
                                     scale0_start = 0.001,
                                     scale0_granularity = 0.05,
                                     scenario = basic100, 
                                     cindex = TRUE)


ggplot(basic100_grid %>%
         filter(Beta %in% c("V1", "V2", "V3", "V4", "V5" ))%>%
         filter(s1 == 1000)
       )+
  geom_point(aes(x=1/s0, y = Estimate, color = Beta))+
  geom_line(aes(x=1/s0, y = Estimate, color = Beta), alpha = 0.5)+
  geom_hline(aes(yintercept = True_beta, color = Beta), linetype = "dashed")+
  geom_hline(aes(yintercept = Sparse_FG_beta, color = Beta), alpha = 0.5)+
  labs(caption = "True estimate in dashed line, sparse fine and gray estimate in solid line",
      title = "Active Coefficients")

ggplot(basic100_grid %>%
         filter(!(Beta %in% c("V1", "V2", "V3", "V4", "V5" )))%>%
         filter(s1 == 1000)
       )+
  geom_point(aes(x=1/s0, y = Estimate, color = Beta))+
  geom_line(aes(x=1/s0, y = Estimate, color = Beta), alpha = 0.5)+
  geom_hline(aes(yintercept = True_beta, color = Beta), linetype = "dashed")+
  geom_hline(aes(yintercept = Sparse_FG_beta, color = Beta), alpha = 0.5)+
  labs(caption = "True estimate in dashed line, sparse fine and gray estimate in solid line",
      title = "Inactive Coefficients")
```

```{r}
target_betas <- c("V1", "V2", "V3", "V4", "V5")

ggplot() +
  # LAYER 1 (BOTTOM): Plot everything EXCEPT V1-V5 in light gray
  geom_line(
    data = subset(basic100_grid, !(Beta %in% target_betas) & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "darkgray"
  ) +
  geom_point(
    data = subset(basic100_grid, !(Beta %in% target_betas)& s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "darkgray"
  ) +
  
  # LAYER 2 (TOP): Plot ONLY V1-V5 in black so they draw over the gray
  geom_line(
    data = subset(basic100_grid, Beta %in% target_betas & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "black"
  ) +
  geom_point(
    data = subset(basic100_grid, Beta %in% target_betas & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "black"
  ) +
  
  # Set the x-axis ticks to match the exact s0 values
  scale_x_continuous(breaks = 1/unique(basic100_grid$s0)) +
  
  # Rotate the text 90 degrees and align it properly
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```

```{r}
basic100_large_grid <- solution_path_wrapper(scale1_seq = c( 1000),
                                     #scale0_path = seq(0.001, 0.1, .005),
                                     scale0_path = 1/(seq(0.0001, 100, 10)),
                                     scale0_start = 0.001,
                                     scale0_granularity = 0.05,
                                     scenario = basic100_large, 
                                     cindex = TRUE)


ggplot(basic100_large_grid %>%
         filter(Beta %in% c("V1", "V2", "V3", "V4", "V5" ))%>%
         filter(s1 == 1000)
       )+
  geom_point(aes(x=1/s0, y = Estimate, color = Beta))+
  geom_line(aes(x=1/s0, y = Estimate, color = Beta), alpha = 0.5)+
  geom_hline(aes(yintercept = True_beta, color = Beta), linetype = "dashed")+
  geom_hline(aes(yintercept = Sparse_FG_beta, color = Beta), alpha = 0.5)+
  labs(caption = "True estimate in dashed line, sparse fine and gray estimate in solid line",
      title = "Active Coefficients")

ggplot(basic100_large_grid %>%
         filter(!(Beta %in% c("V1", "V2", "V3", "V4", "V5" )))%>%
         filter(s1 == 1000)
       )+
  geom_point(aes(x=1/s0, y = Estimate, color = Beta))+
  geom_line(aes(x=1/s0, y = Estimate, color = Beta), alpha = 0.5)+
  geom_hline(aes(yintercept = True_beta, color = Beta), linetype = "dashed")+
  geom_hline(aes(yintercept = Sparse_FG_beta, color = Beta), alpha = 0.5)+
  labs(caption = "True estimate in dashed line, sparse fine and gray estimate in solid line",
      title = "Inactive Coefficients")
```

```{r}
ggplot() +
  # LAYER 1 (BOTTOM): Plot everything EXCEPT V1-V5 in light gray
  geom_line(
    data = subset(basic100_large_grid, !(Beta %in% target_betas) & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "darkgray"
  ) +
  geom_point(
    data = subset(basic100_large_grid, !(Beta %in% target_betas)& s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "darkgray"
  ) +
  
  # LAYER 2 (TOP): Plot ONLY V1-V5 in black so they draw over the gray
  geom_line(
    data = subset(basic100_large_grid, Beta %in% target_betas & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "black"
  ) +
  geom_point(
    data = subset(basic100_large_grid, Beta %in% target_betas & s1 == 1000),
    aes(x = 1/s0, y = Estimate, group = Beta),
    color = "black"
  ) +
  
  # Set the x-axis ticks to match the exact s0 values
  scale_x_continuous(breaks = 1/unique(basic100_large_grid$s0)) +
  
  # Rotate the text 90 degrees and align it properly
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)
  )
```
