---
title: "Untitled"
format: html
editor: visual
---

# Set up

```{r}
library(tidyverse)
library(fastcmprsk, lib.loc = "/Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library")
library(survival)
Rcpp::compileAttributes("~/Documents/bhCRR")
devtools::load_all("~/Documents/bhCRR", recompile = TRUE)
```

```{r}


 # Block-correlated design matrix generator (Binder & Schumacher 2008 sec 3.1;
 # Binder et al. 2009 sec 3.2.1). Kept in its own file so batch_run_complex.R
 # uses the identical generator instead of a drifting copy.
 source("/Users/sophiehuebler/Documents/bhCRR/Sim/code/generate_block_X.R")



#################### Data Generation ################

# To change per scenario
nobs = 200
npredictors = 25

# Effects of the designed active predictors. beta1_active, beta2_active and
# active_block are aligned elementwise: entry k is one predictor, with its
# cause-1 effect, its cause-2 effect, and the block it is planted in. An entry
# with one beta set to 0 is still a designed active, which is how Binder et al.
# (2009) place covariates acting on a single cause.
beta1_active = c(0.40, - 0.50, 0.60, 0.75, - 0.80)
beta2_active = c(0, 0.3, 0, 0, -0.2)
active_block = c("strong", "strong", "weak", "weak", "indep")

# Predictor correlation blocks. block_props are relative shares of npredictors
# (apportioned by largest remainder so they sum exactly); block_rho is the
# within-block correlation, which is exchangeable inside a block and exactly
# zero between blocks. Names must match across all three vectors.
block_props = c(indep = 0.50, weak = 0.25, strong = 0.25)
block_rho   = c(indep = 0.00, weak = 0.30, strong = 0.60)

# Latent driving the within-block correlation. "gaussian" hits block_rho exactly
# in every sample. "bernoulli" is the papers' own construction, giving bimodal
# microarray-like marginals, and then latent_q sets the shift probability per
# block. To replicate Binder et al. (2009) directly, use latent = "bernoulli"
# with block_rho = c(indep = 0, weak = 0.05, strong = 0.5) and
# latent_q = c(indep = 0.5, weak = 0.7, strong = 0.5).
latent_type = "gaussian"
latent_q    = c(indep = 0.5, weak = 0.4, strong = 0.5)

# Random censoring. Censoring times are drawn as C_i ~ Exponential(cens_rate),
# independently of the predictors and of the event process. A larger rate means
# censoring happens earlier, so a larger rate gives a higher censored fraction.
# The mean censoring time is 1 / cens_rate, so pick the rate relative to the
# scale of the event times (roughly, rate ~ 1 / (typical follow-up time)).
# Set cens_rate = 0 to turn random censoring off entirely.
cens_rate = 0.05

# Administrative censoring bounds handed to simulateTwoCauseFineGrayModel().
# Keeping u_min = u_max = 100 leaves the generator's own uniform censoring
# inactive (no event time reaches 100), so cens_rate is the only censoring
# mechanism. Lower these if administrative censoring is wanted as well.
u_min = 100
u_max = 100


# Calculated (fixed)
# generate_block_X() builds the block-correlated design matrix, places each
# designed active in the block named for it, and scatters the betas to the
# matching columns. With layout = "active_first" the actives occupy columns
# 1..length(beta1_active) in the order written above, so beta1[k] is the effect
# of beta1_active[k]. design$meta carries the per-predictor block labels.
design <- generate_block_X(nobs          = nobs,
                           npredictors   = npredictors,
                           beta1_active  = beta1_active,
                           beta2_active  = beta2_active,
                           active_block  = active_block,
                           block_props   = block_props,
                           block_rho     = block_rho,
                           latent        = latent_type,
                           latent_q      = latent_q,
                           unit_variance = TRUE,
                           layout        = "active_first")

beta1 <- design$beta1
beta2 <- design$beta2
X_design <- design$X

# Realized block structure for this replicate. Worth eyeballing when a new
# block specification is used for the first time.
# print(design$block_size)
# print(block_corr_summary(X_design, design$block_id, design$block_rho))

# Simulated
sim <- simulateTwoCauseFineGrayModel(nobs = nobs,
                                     beta1 = beta1,
                                     beta2 = beta2,
                                     X = X_design,
                                     u.min = u_min,
                                     u.max = u_max,
                                     p = 0.5,
                                     returnX = TRUE)

# The event times are only interpretable if the generator used the design matrix
# we handed it rather than one of its own. This catches a silent substitution.
# It cannot catch internal rescaling of X prior to generating the event times,
# which is worth confirming once against the fastcmprsk source.
if (!isTRUE(all.equal(unname(as.matrix(sim$X)), unname(X_design),
                      tolerance = 1e-12))) {
  stop("simulateTwoCauseFineGrayModel() did not return the supplied design ",
       "matrix unchanged; the event times were not generated from X_design.")
}

# Apply the exponential random censoring on top of the generated event times.
# Subject i is censored when C_i lands before the event time already simulated
# for them, in which case the observed time becomes C_i and Status becomes 0.
# Overwriting sim$ftime / sim$fstatus here (rather than only sim_data) keeps
# every downstream consumer of `sim`, including the oracle Fine-Gray fit, on the
# same censored data.
if (!is.finite(cens_rate) || cens_rate < 0) {
  stop("cens_rate must be a finite, non-negative Exponential rate.")
}

if (cens_rate > 0) {
  cens_time <- stats::rexp(nobs, rate = cens_rate)
  is_censored <- cens_time < sim$ftime
  sim$ftime[is_censored] <- cens_time[is_censored]
  sim$fstatus[is_censored] <- 0
}

# Realized outcome mix for this replicate. Worth logging per scenario: the
# censored fraction is a consequence of cens_rate and the event-time scale, not
# something set directly.
cens_prop <- mean(sim$fstatus == 0)
event_mix <- c(censored = cens_prop,
               cause1 = mean(sim$fstatus == 1),
               cause2 = mean(sim$fstatus == 2))

sim_data <- cbind(data.frame(ID = 1:nobs,
                             TTE = sim$ftime,
                             Status = as.integer(sim$fstatus)),
                  X_design)
predictor_names <- design$meta$name
names(sim_data) <- c("ID", "TTE", "Status", predictor_names)

x <- as.matrix(sim_data %>%select(starts_with("X")))
y <- as.matrix(sim_data %>%
                 mutate(Status = as.numeric(Status))%>%
                 select(TTE, Status))

storage.mode(x) <- "double"

if (!all(sim_data$Status %in% c(0L, 1L, 2L))) {
  stop("Status must use 0 = censored, 1 = event of interest, 2 = competing event.")
}
if (sum(sim_data$Status == 1L) == 0L) {
  stop("The simulated data contain no cause-1 events; change the seed or scenario.")
}
if (cens_rate > 0 && cens_prop > 0.9) {
  warning("Over 90% of observations are censored at cens_rate = ", cens_rate,
          "; the scenario may be uninformative.")
}

# Standardize ONCE and feed this same matrix to every method. This avoids
# method-to-method differences from hidden preprocessing, and it puts the
# adaptive weights on the same coefficient scale as the penalized fits.
x_center <- colMeans(x)
x_scale <- apply(x, 2L, stats::sd)

bad_scale <- !is.finite(x_scale) | x_scale <= sqrt(.Machine$double.eps)
if (any(bad_scale)) {
  stop(
    "Predictors with zero or invalid standard deviation: ",
    paste(predictor_names[bad_scale], collapse = ", ")
  )
}

x_std <- sweep(x, 2L, x_center, FUN = "-")
x_std <- sweep(x_std, 2L, x_scale, FUN = "/")
colnames(x_std) <- predictor_names

fit_data <- data.frame(
  TTE = sim_data$TTE,
  Status = sim_data$Status,
  x_std,
  check.names = FALSE
)

fg_formula <- stats::as.formula(
  paste0(
    "Crisk(TTE, Status, cencode = 0, failcode = 1) ~ ",
    paste(predictor_names, collapse = " + ")
  )
)


#################### FG Model ################
active_idx <- which(beta1 != 0)
active_x <- x[, active_idx, drop = FALSE]

fg_true <- fastcmprsk::fastCrr(fastcmprsk::Crisk(sim$ftime,
                                                 as.integer(sim$fstatus)) ~ active_x,
                               variance = FALSE)




```

# Testing

## 1. Untuned fit with the BIC initialization

`fit_ssl_psdh()` now requires an explicit initialization: either `init` (a length-p numeric vector) or `init_method`. `"LASSO_bic"` fits a single `fastCrrp` LASSO path and takes the coefficients at the lowest-BIC lambda among the non-null, converged solutions -- one path fit, versus the six that `"LASSO_cv"` needs (one full-data fit plus one per CV fold).

```{r}
untuned <- fit_ssl_psdh(
  x_std, y,
  ss          = c(0.04, 0.6),
  init_method = "LASSO_bic",
  # init_args overrides the method's own defaults; drop it to use
  # .default_init_lam_path(), which is this same sequence.
  init_args   = list(lambda_path = 10^seq(log10(0.1), log10(0.001),
                                          length.out = 10))
)

# Provenance for the initialization now travels on the fit object.
untuned$init_method                       # "LASSO_bic"
untuned$init_meta[c("lambda", "bic", "df")]
sum(untuned$init != 0)                    # nonzero coefficients in the init

# The fit itself
untuned$conv                              # outer EM convergence
untuned$iterations
sum(untuned$coefficients$Estimate != 0)   # nonzero coefficients after EM
head(untuned$coefficients[order(-abs(untuned$coefficients$Estimate)), ], 10)

# Equivalent alternatives:
#   init_method = "LASSO_cv"                     # the old built-in behaviour
#   init_method = "zero"                         # rep(0, p), no initial fit
#   init        = rep(0, ncol(x_std))            # supply a vector directly
#   init_method = function(x, y, ...) rep(0, ncol(x))   # your own function
# Supplying both init and init_method is an error, by design.
```

## 2. Fixed-grid cross-validation

Two new pieces are in place: `bhcrr_cv_control()` bundles every CV setting into one validated object, and `bhcrr_make_folds()` builds stratified folds through `rsample`. Neither is wired into `tune_ssl_psdh()` yet -- that is the CV rewrite -- so below the control's fields are handed across by name, which is a fair preview of what the rewritten signature will absorb.

Two things to know before running this:

-   **`warm_start` must be `FALSE` for now.** With warm starts on, the first `(s0, s1)` pair passes `init = NULL` and the rest pass a fitted vector, which collides with `init_method` and errors on every pair after the first. The error is swallowed by `try()` inside `cv_ssl_psdh()`, so it shows up as `NA` scores rather than a stop. Turning warm starts off makes every fold cold-start from `init_method`, which is correct, just slower.
-   **`pool = 0.01`, not rsample's default `0.1`.** Any stratum below `pool` is pooled away and randomly reassigned, so at a cause-1 rate near 10% the default would silently give unstratified folds.

```{r}
ctrl <- bhcrr_cv_control(
  nfolds        = 5,            # production runs use 10
  ncv           = 2,            # production runs use 2
  strata        = "cause1",     # balance cause-1 events across folds
  pool          = 0.01,
  eval_quantile = 0.5,
  init_method   = "LASSO_bic",
  warm_start    = FALSE,        # required until the CV rewrite; see above
  # fit_args is stored but CANNOT be forwarded to tune_ssl_psdh() today:
  # cv_ssl_psdh() hardcodes maxit = 50 and epsilon = 1e-04 in its own call to
  # fit_ssl_psdh(), so passing either through ... errors with "formal argument
  # matched by multiple actual arguments". Same collision class as `init`.
  fit_args      = list(maxit = 50, epsilon = 1e-04)
)
ctrl

# Stratified folds. fold_event_counts is the per-fold cause-1 count: with
# strata = "cause1" these should differ by at most 1.
folds <- bhcrr_make_folds(y, ctrl)
folds$fold_event_counts

# Fixed grid: s0 in [0.01, 0.10], s1 in [0.40, 0.70]. Pairs with s1 <= s0 are
# dropped internally; here every combination is valid, so 5 x 4 = 20 pairs.
s0_seq <- seq(0.01, 0.10, length.out = 5)
s1_seq <- seq(0.40, 0.70, length.out = 4)

# tune_ssl_psdh() uses `untuned` only as a data container ($x and $y) -- one of
# the things the rewrite will drop.
t0 <- Sys.time()
tuning <- tune_ssl_psdh(
  untuned,
  s0_seq      = s0_seq,
  s1_seq      = s1_seq,
  nfolds      = ctrl$nfolds,
  ncv         = ctrl$ncv,
  foldid      = folds$foldid,
  warm_start    = ctrl$warm_start,
  eval_quantile = ctrl$eval_quantile,  # matches a cv_ssl_psdh formal
  init_method   = ctrl$init_method,    # reaches fit_ssl_psdh through ...
  init_args     = ctrl$init_args
)
Sys.time() - t0

# Check for the silent failure mode FIRST: NA scores mean the fits errored
# inside try() rather than that the pairs were bad.
sum(is.na(tuning$score_mean))
sum(tuning$n_failed_folds, na.rm = TRUE)

tuning[order(-tuning$score_mean), ][1:5, ]
best <- tuning[which.max(tuning$score_mean), ]
best

# The grid diagnostics still work, since they only read the tuning frame plus
# theta from the untuned fit.
theta <- mean(untuned$coefficients$Estimate != 0)
bhcrr_tune_validate(tuning, n = nrow(x_std), theta = theta,
                    s0_seq = s0_seq, s1_seq = s1_seq)
```

## 3. Refit at the tuned scales

```{r}
final_mod <- fit_ssl_psdh(
  x_std, y,
  ss          = c(best$s0, best$s1),
  init_method = "LASSO_bic",
  maxit       = 100
)

final_mod$conv
sum(final_mod$coefficients$Estimate != 0)

# Breslow jumps are now computed once, in a single refit after the EM loop,
# so predict_from_ssl_psdh() still has what it needs.
eval_time <- as.numeric(quantile(y[y[, 2] == 1, 1], 0.5))
lp <- predict_from_ssl_psdh(final_mod, newx = x_std, eval_time)
wolbers_c(y, lp, eval_time1)
```

# Testing Serial

The Prompt C checkpoint: run `bhcrr_cv()` end to end on this simulated data with a real grid, sequentially, before any parallelism exists. A wrong assembler is far easier to spot here than after `mclapply` is in the picture.

### C.1 — Set up and run

Two deliberate choices:

-   **Folds are built once and passed to every call.** If `folds = NULL`, `bhcrr_cv()` generates them internally, and unless `control$seed` is set they differ between calls — which would make the `fold_inits` round trip in C.3 fail for a reason that has nothing to do with the caching.
-   **`eval_time` is supplied explicitly**, so tau is a fixed a-priori horizon rather than something estimated from `y`. `$eval_time$source` should read `"supplied"`.

```{r}
# Fixed horizon. Computed here from the observed cause-1 times purely to land on a
# sensible number for this simulation; once chosen it is a constant, not an estimate.
tau <- as.numeric(quantile(y[y[, 2] == 1, 1], 0.5))

ctrl_serial <- bhcrr_cv_control(
  nfolds      = 5,
  ncv         = 2,
  strata      = "cause1",
  pool        = 0.01,
  eval_time   = tau,
  init_method = "LASSO_bic",
  warm_start  = TRUE,
  parallel    = FALSE,
  keep_coefs  = FALSE
)
ctrl_serial

# Build folds ONCE and reuse. fold_event_counts should differ by at most 1 per fold.
folds_serial <- bhcrr_make_folds(y, ctrl_serial)
folds_serial$fold_event_counts

s0_seq <- seq(0.01, 0.10, length.out = 5)
s1_seq <- seq(0.40, 0.70, length.out = 4)

t0 <- Sys.time()
cv <- bhcrr_cv(x_std, y,
               s0_seq = s0_seq,
               s1_seq = s1_seq,
               control = ctrl_serial,
               folds   = folds_serial)
elapsed_serial <- Sys.time() - t0
elapsed_serial

cv                      # print.bhcrr_cv
class(cv)
cv$eval_time            # $value == tau, $source == "supplied"
cv$timing
```

### C.2 — Assembler consistency

This is the actual checkpoint. Each block re-derives a column of `$tuning` from the raw `$pooled` / `$fold_scores` arrays and asserts they agree. If the assembler is wrong, one of these fails loudly rather than returning a plausible-looking frame.

```{r}
n_pairs <- nrow(cv$grid)
n_folds <- ctrl_serial$nfolds
n_reps  <- ctrl_serial$ncv

# --- structure -------------------------------------------------------------
stopifnot(
  nrow(cv$tuning) == n_pairs,
  identical(dim(cv$pooled),      c(n_pairs, n_reps)),
  identical(dim(cv$fold_scores), c(n_pairs, n_folds, n_reps))
)

# $tuning must carry exactly the columns bhcrr_tune_validate() expects
stopifnot(all(c("s0", "s1", "score_mean", "score_sd",
                "score_fold_sd", "n_failed_folds") %in% names(cv$tuning)))

# --- pair ordering ---------------------------------------------------------
# $tuning rows must be in .cv_grid traversal order, which is the warm-start
# chain order. If these disagree, scores are attached to the wrong pairs.
stopifnot(
  isTRUE(all.equal(cv$tuning$s0, cv$grid$s0)),
  isTRUE(all.equal(cv$tuning$s1, cv$grid$s1))
)

# --- score_mean is the mean of the pooled scores over reps ------------------
expected_mean <- rowMeans(cv$pooled, na.rm = TRUE)
stopifnot(isTRUE(all.equal(cv$tuning$score_mean, expected_mean,
                           tolerance = 1e-12)))

# --- score_sd is the sd of the pooled scores over reps ----------------------
# NA when ncv == 1; with ncv = 2 it is the sd of two numbers.
expected_sd <- if (n_reps == 1L) rep(NA_real_, n_pairs) else
  apply(cv$pooled, 1, stats::sd, na.rm = TRUE)
stopifnot(isTRUE(all.equal(cv$tuning$score_sd, expected_sd, tolerance = 1e-12)))

# --- score_fold_sd is the mean over reps of the within-rep fold sd ----------
expected_fold_sd <- vapply(seq_len(n_pairs), function(j) {
  per_rep <- vapply(seq_len(n_reps),
                    function(k) stats::sd(cv$fold_scores[j, , k], na.rm = TRUE),
                    numeric(1))
  mean(per_rep, na.rm = TRUE)
}, numeric(1))
stopifnot(isTRUE(all.equal(cv$tuning$score_fold_sd, expected_fold_sd,
                           tolerance = 1e-12)))

# --- failures ---------------------------------------------------------------
# n_failed_folds must account for every row in $errors, and vice versa.
stopifnot(sum(cv$tuning$n_failed_folds) == nrow(cv$errors))
nrow(cv$errors)
if (nrow(cv$errors) > 0) head(cv$errors)

# --- inits ------------------------------------------------------------------
# One init per (rep, fold), each length p.
stopifnot(
  length(cv$fold_inits) == n_reps,
  all(lengths(cv$fold_inits) == n_folds),
  all(vapply(unlist(cv$fold_inits, recursive = FALSE), length, integer(1)) ==
        ncol(x_std))
)

# --- sanity on the values ---------------------------------------------------
# A C-index lives in [0, 1]; anything outside means the scatter of held-out
# linear predictors into lp_all went wrong.
finite_scores <- cv$tuning$score_mean[is.finite(cv$tuning$score_mean)]
stopifnot(length(finite_scores) > 0,
          all(finite_scores >= 0), all(finite_scores <= 1))
range(finite_scores)

cat("assembler checks passed\n")
```

If any `all.equal` check fails, look first at whether the assembler and these re-derivations disagree on `na.rm` rather than assuming a real bug — the old `cv_ssl_psdh` used `na.rm = TRUE` throughout, and that convention is what is assumed above.

### C.3 — `fold_inits` round trip

The second run reuses the cached inits and the same folds, so it must reproduce `$tuning` exactly while skipping every initialization fit. The time difference is the init cost the cache removes.

```{r}
t0 <- Sys.time()
cv2 <- bhcrr_cv(x_std, y,
                s0_seq     = s0_seq,
                s1_seq     = s1_seq,
                control    = ctrl_serial,
                folds      = folds_serial,
                fold_inits = cv$fold_inits)
elapsed_cached <- Sys.time() - t0

stopifnot(isTRUE(all.equal(cv$tuning, cv2$tuning, tolerance = 1e-12)))

c(first = as.numeric(elapsed_serial, units = "secs"),
  cached = as.numeric(elapsed_cached, units = "secs"))
```

### C.4 — Derived horizon, for contrast

Leaving `eval_time = NULL` falls back to the `eval_quantile` of the cause-1 times in the full data. `$eval_time$source` should now read `"quantile"`. Scores will differ slightly from C.1 whenever the derived tau differs from the supplied one.

```{r}
ctrl_derived <- bhcrr_cv_control(
  nfolds        = 5,
  ncv           = 2,
  strata        = "cause1",
  pool          = 0.01,
  eval_time     = NULL,
  eval_quantile = 0.5,
  init_method   = "LASSO_bic",
  warm_start    = TRUE,
  parallel      = FALSE
)

cv_derived <- bhcrr_cv(x_std, y, s0_seq, s1_seq,
                       control    = ctrl_derived,
                       folds      = folds_serial,
                       fold_inits = cv$fold_inits)

cv_derived$eval_time          # $source == "quantile"
c(supplied = cv$eval_time$value, derived = cv_derived$eval_time$value)
```

### C.5 — Diagnostics and the final refit

`$tuning` must drop straight into the existing grid diagnostics, unchanged.

```{r}
cv$best

theta <- mean(untuned$coefficients$Estimate != 0)
bhcrr_tune_validate(cv$tuning,
                    n      = nrow(x_std),
                    theta  = theta,
                    s0_seq = s0_seq,
                    s1_seq = s1_seq)

# Refit at the winner and score it at the same fixed horizon.
final_serial <- fit_ssl_psdh(
  x_std, y,
  ss          = c(cv$best$s0, cv$best$s1),
  init_method = "LASSO_bic",
  maxit       = 100
)

final_serial$conv
sum(final_serial$coefficients$Estimate != 0)

lp_final <- predict_from_ssl_psdh(final_serial, newx = x_std, tau)
wolbers_c(y, lp_final, tau)
```

Note this last number is an **in-sample** C-index — the model was fit on all of `x_std`. It is not comparable to `cv$best$score_mean`, which is out-of-sample. It is here only to confirm the refit produces a usable model with working Breslow jumps.

### C.6 — Convergence diagnostics

Non-convergence is a `warning()` in `fit_ssl_psdh()`, not an error: the fit still returns a usable object, so `.cv_fold_path()`'s `tryCatch(error = )` never sees it and `$errors` stays empty. That is correct behaviour, but it means non-converged fits are scored and averaged into `$tuning` with nothing marking them.

Note the denominator: this run does `nfolds * ncv * npairs` fits, not one per pair.

```{r}
n_fits_total <- ctrl_serial$nfolds * ctrl_serial$ncv * nrow(cv$grid)
n_fits_total
```

#### Count the warnings exactly

`withCallingHandlers` captures each warning as it is signalled, so nothing is lost to R's end-of-call batching or the 50-warning display cap.

```{r}
warn_log <- character(0)

cv_dbg <- withCallingHandlers(
  bhcrr_cv(x_std, y,
           s0_seq     = s0_seq,
           s1_seq     = s1_seq,
           control    = ctrl_serial,
           folds      = folds_serial,
           fold_inits = cv$fold_inits),   # reuse cached inits: this is about the EM
  warning = function(w) {
    warn_log <<- c(warn_log, conditionMessage(w))
    invokeRestart("muffleWarning")
  }
)

length(warn_log)                          # total warnings signalled
length(warn_log) / n_fits_total           # as a fraction of all fits
table(warn_log)                           # distinct messages, with counts
```

#### Is it the outer EM or the inner solver?

The same warning text fires for two different failures. `fit_ssl_psdh()` breaks out of the EM loop *before* incrementing `iterations`, so:

-   `iterations == maxit` → the outer EM ran its full budget without meeting the deviance-plus-stability test. Loosen `epsilon` or raise `maxit`.
-   `iterations < maxit` **and** `conv == FALSE` → the inner `fastCrrp` solver hit its escalation ceiling (`inner_maxit` climbs to 20000) and the EM stopped early. This is the more serious of the two: the M-step estimate at that point is untrustworthy.

Walk one fold's chain by hand, recording what `bhcrr_cv` currently discards. This replicates `.cv_fold_path()` exactly — same init, same traversal order, same warm-start chain — but keeps `$conv` and `$iterations` for every pair.

```{r}
maxit_used <- if (!is.null(ctrl_serial$fit_args$maxit)) ctrl_serial$fit_args$maxit else 50L

tr   <- which(folds_serial$foldid[, 1] != 1L)   # rep 1, fold 1 training rows
x_tr <- x_std[tr, , drop = FALSE]
y_tr <- y[tr, , drop = FALSE]

# Same init the worker would compute for this fold
b <- bhCRR:::.LASSO_bic(x_tr, y_tr)$init

diag_rows <- vector("list", nrow(cv$grid))
for (j in seq_len(nrow(cv$grid))) {
  f <- suppressWarnings(
    fit_ssl_psdh(x_tr, y_tr,
                 ss    = c(cv$grid$s0[j], cv$grid$s1[j]),
                 init  = b,
                 maxit = maxit_used)
  )
  diag_rows[[j]] <- data.frame(
    pair       = cv$grid$pair[j],
    s0         = cv$grid$s0[j],
    s1         = cv$grid$s1[j],
    conv       = f$conv,
    iterations = f$iterations,
    nonzero    = sum(f$coefficients$Estimate != 0)
  )
  b <- as.numeric(f$final_model_object$coef)   # warm-start chain
}
conv_diag <- do.call(rbind, diag_rows)
conv_diag

# Classify the failures
failed <- conv_diag[!conv_diag$conv, ]
cat("not converged:", nrow(failed), "of", nrow(conv_diag), "pairs on this fold\n")
if (nrow(failed) > 0) {
  cat("  outer EM exhausted (iterations == maxit):",
      sum(failed$iterations == maxit_used), "\n")
  cat("  inner solver broke early (iterations < maxit):",
      sum(failed$iterations < maxit_used), "\n")
  print(failed)
}
```

#### Where in the grid does it fail?

If non-convergence clusters at one end of the grid — most likely the small-`s0` end, where the spike is sharpest and the penalty surface hardest — that is informative rather than alarming, and argues for trimming the grid rather than raising `maxit`.

```{r}
with(conv_diag, table(s0 = s0, converged = conv))
with(conv_diag, table(s1 = s1, converged = conv))
```

#### If it is the outer EM

Confirm by giving it more room on one of the failing pairs. If `maxit = 200` converges, it was a budget problem; if it still exhausts, the deviance criterion is not settling and `epsilon` is the thing to look at.

```{r}
if (nrow(failed) > 0) {
  worst <- failed[1, ]
  f200 <- suppressWarnings(
    fit_ssl_psdh(x_tr, y_tr, ss = c(worst$s0, worst$s1),
                 init_method = "LASSO_bic", maxit = 200L)
  )
  c(s0 = worst$s0, s1 = worst$s1,
    conv_at_50 = worst$conv, iters_at_50 = worst$iterations,
    conv_at_200 = f200$conv, iters_at_200 = f200$iterations)
}
```

#### Reconcile across all chains

The by-hand walk above covers ONE of the `nfolds * ncv` chains — 20 of the 200 fits. Zero failures there is consistent with 22 failures overall; it just means this chain was not one of the affected ones.

Sweep every chain and check the total comes to exactly `length(warn_log)`. If it does, the warnings are fully explained and there is nothing further to chase. If it does not, something is signalling warnings that is not the EM, and the message table below will say what.

Each chain starts from the exact init the run used (`cv$fold_inits[[k]][[i]]`), so this reproduces the fits rather than approximating them. Cost is the same 200 fits as C.1.

```{r}
sweep_rows <- vector("list", 0)

for (k in seq_len(ctrl_serial$ncv)) {
  for (i in seq_len(ctrl_serial$nfolds)) {

    tr   <- which(folds_serial$foldid[, k] != i)
    x_tr <- x_std[tr, , drop = FALSE]
    y_tr <- y[tr, , drop = FALSE]

    b <- cv$fold_inits[[k]][[i]]        # the exact init this chain used

    for (j in seq_len(nrow(cv$grid))) {
      f <- suppressWarnings(
        fit_ssl_psdh(x_tr, y_tr,
                     ss    = c(cv$grid$s0[j], cv$grid$s1[j]),
                     init  = b,
                     maxit = maxit_used)
      )
      sweep_rows[[length(sweep_rows) + 1L]] <- data.frame(
        rep = k, fold = i, pair = cv$grid$pair[j],
        s0 = cv$grid$s0[j], s1 = cv$grid$s1[j],
        conv = f$conv, iterations = f$iterations
      )
      b <- as.numeric(f$final_model_object$coef)
    }
  }
}
conv_sweep <- do.call(rbind, sweep_rows)

# --- the reconciliation ----------------------------------------------------
n_not_conv <- sum(!conv_sweep$conv)
c(fits = nrow(conv_sweep),
  not_converged = n_not_conv,
  warnings_captured = length(warn_log))

if (n_not_conv == length(warn_log)) {
  cat("reconciled: every warning is one non-converged EM fit\n")
} else {
  cat("NOT reconciled — inspect the distinct warning messages:\n")
  print(table(warn_log))
}
```

```{r}
# --- which chains? ---------------------------------------------------------
# Clustering by (rep, fold) means specific training subsets are hard, not that
# the grid is bad.
with(conv_sweep, table(rep = rep, fold = fold, converged = conv))

# --- where in the grid? ----------------------------------------------------
with(conv_sweep, table(s0 = s0, converged = conv))
with(conv_sweep, table(s1 = s1, converged = conv))

# --- outer EM or inner solver? ---------------------------------------------
failed_sweep <- conv_sweep[!conv_sweep$conv, ]
if (nrow(failed_sweep) > 0) {
  cat("outer EM exhausted (iterations == maxit):",
      sum(failed_sweep$iterations == maxit_used), "\n")
  cat("inner solver broke early (iterations < maxit):",
      sum(failed_sweep$iterations <  maxit_used), "\n")
  failed_sweep[order(failed_sweep$rep, failed_sweep$fold, failed_sweep$pair), ]
}

# --- does it move the scores? ----------------------------------------------
# Count non-converged fits per pair and set it beside that pair's CV score. If
# the affected pairs are the ones losing anyway, this changes nothing; if they
# are near the top of the grid, the winner is being chosen partly on
# untrustworthy fits.
per_pair <- aggregate(cbind(n_not_conv = !conv) ~ pair, data = conv_sweep, FUN = sum)
score_vs_conv <- merge(
  cbind(cv$tuning, pair = cv$grid$pair),
  per_pair, by = "pair"
)
score_vs_conv[order(-score_vs_conv$score_mean),
              c("pair", "s0", "s1", "score_mean", "n_not_conv")]
```

#### Does non-convergence change the winner?

Confirmed by reading `R/cv_fold_path.r` and `R/cv.r`: a non-converged fit takes the success branch (`cv_fold_path.r:85`), its linear predictor is stored exactly like a converged one (`:97`), and it flows into both the per-fold and pooled C-index (`cv.r:198`, `:215`) and then into `score_mean` (`:223`). Nothing in `cv.r` references `conv`. It also seeds the next pair's warm start (`:105`), so its influence is not confined to the pairs that warned.

That is not automatically wrong — if the estimator is "EM with at most `maxit` iterations", then a non-converged fit *is* what the procedure produces, and dropping those fits would bias the score upward for exactly the pairs where convergence is hard. The question is whether it changes the ranking.

Mask the non-converged fits and recompute. This uses the mean of per-fold scores rather than the pooled score (the pooled version would need the held-out linear predictors, which `bhcrr_cv()` does not return), so treat it as a ranking sensitivity check, not an exact recomputation of `score_mean`.

```{r}
# conv indicator aligned to cv$fold_scores: npairs x nfolds x ncv
conv_arr <- array(NA, dim = dim(cv$fold_scores))
for (r in seq_len(nrow(conv_sweep))) {
  conv_arr[conv_sweep$pair[r], conv_sweep$fold[r], conv_sweep$rep[r]] <-
    conv_sweep$conv[r]
}
stopifnot(!anyNA(conv_arr))          # every cell accounted for

fs_all  <- cv$fold_scores
fs_conv <- fs_all
fs_conv[!conv_arr] <- NA_real_       # drop non-converged fits

cmp <- data.frame(
  pair        = cv$grid$pair,
  s0          = cv$grid$s0,
  s1          = cv$grid$s1,
  score_mean  = cv$tuning$score_mean,                    # as reported
  mean_all    = apply(fs_all,  1, mean, na.rm = TRUE),   # all fits
  mean_conv   = apply(fs_conv, 1, mean, na.rm = TRUE),   # converged only
  n_fits      = apply(!is.na(fs_all),  1, sum),
  n_converged = apply(!is.na(fs_conv), 1, sum)
)
cmp$n_not_conv <- cmp$n_fits - cmp$n_converged
cmp$delta      <- cmp$mean_conv - cmp$mean_all

cmp[order(-cmp$score_mean), ]
```

```{r}
# --- does the argmax move? -------------------------------------------------
win_reported <- cmp$pair[which.max(cmp$score_mean)]
win_all      <- cmp$pair[which.max(cmp$mean_all)]
win_conv     <- cmp$pair[which.max(cmp$mean_conv)]

c(reported = win_reported, fold_mean_all = win_all, fold_mean_converged = win_conv)

if (win_conv == win_reported) {
  cat("winner is unchanged when non-converged fits are dropped\n")
} else {
  cat("WINNER MOVES: pair", win_reported, "->", win_conv, "\n")
  cmp[cmp$pair %in% c(win_reported, win_conv), ]
}

# --- how exposed is the reported winner? -----------------------------------
cmp[cmp$pair == win_reported, c("pair","s0","s1","score_mean",
                                "n_fits","n_not_conv","mean_all","mean_conv","delta")]

# --- is the margin bigger than the convergence effect? ---------------------
# If the gap between the top two pairs is smaller than |delta| at the winner,
# the ranking is not robust to how non-convergence is handled.
top2   <- sort(cmp$score_mean, decreasing = TRUE)[1:2]
margin <- top2[1] - top2[2]
c(margin_top2 = margin,
  delta_at_winner = cmp$delta[cmp$pair == win_reported],
  robust = abs(cmp$delta[cmp$pair == win_reported]) < margin)
```

```{r}
# --- budget mismatch between CV and the final refit ------------------------
# The CV fits ran at the fit_ssl_psdh default maxit; C.5 refits the winner at
# maxit = 100. If the winner's CV fits were hitting the cap, the model that was
# TUNED is not the model that gets FIT.
c(maxit_in_cv    = maxit_used,
  maxit_in_refit = 100L,
  winner_non_converged_fits = cmp$n_not_conv[cmp$pair == win_reported])
```

# Parallel with Warm Start

# Testing Parallel

Warm starts won the benchmark, so this is the configuration to use: `parallel = TRUE` with `warm_start = TRUE`, which dispatches one task per `(rep, fold)` chain. Width is `nfolds * ncv` — the chain is sequential along the grid by construction, so the pair dimension cannot be split while warm starts are on.

> **Running this in Positron.** Forking is unreliable inside RStudio and Positron and can hang the session; `bhcrr_cv()` emits a message about it rather than silently falling back. If the chunk stalls, interrupt it and either set `parallel = FALSE`, or run the file with `quarto render` / `Rscript`, where `interactive()` is `FALSE` and forking is well behaved.

### P.1 — Run the tuning grid in parallel

Two practices this demonstrates deliberately:

-   **`maxit` is set explicitly in `fit_args`, and the final refit in P.5 uses the same value.** Relying on the `fit_ssl_psdh()` default for CV and then refitting at a different budget means the model that was tuned is not the model that gets fit.
-   **`eval_time` is supplied**, so tau is a fixed a-priori horizon rather than something estimated from `y`.

```{r}
s0_seq <- seq(0.01, 0.08, length.out = 7)
s1_seq <- seq(0.40, 0.70, length.out = 4)   # 7 x 4 = 28 pairs, all valid

maxit_cv <- 50L    # used for the CV fits AND the final refit in P.5

ctrl_par <- bhcrr_cv_control(
  nfolds      = 10,
  ncv         = 2,
  strata      = "cause1",
  pool        = 0.01,
  eval_time   = tau,
  init_method = "LASSO_bic",
  warm_start  = TRUE,        # -> dispatch = "fold", width nfolds * ncv = 20
  fit_args    = list(maxit = maxit_cv, epsilon = 1e-04),
  parallel    = TRUE,
  workers     = max(1L, parallel::detectCores() - 1L)
)
ctrl_par

folds_par <- bhcrr_make_folds(y, ctrl_par)
folds_par$fold_event_counts

t0 <- Sys.time()
cv_par <- bhcrr_cv(x_std, y,
                   s0_seq  = s0_seq,
                   s1_seq  = s1_seq,
                   control = ctrl_par,
                   folds   = folds_par)
elapsed_par <- Sys.time() - t0

cv_par
cv_par$timing            # $dispatch should be "fold"; $workers as requested
nrow(cv_par$grid)        # 28 pairs
prod(dim(cv_par$fold_scores))   # total fits = 28 * 10 * 2
```

### P.2 — Parallel must equal serial

The chains are independent and every input is fixed — folds are passed in, inits are cached and reused, traversal order is deterministic — so the parallel run must reproduce the serial one exactly. Anything else means non-determinism leaked into a worker.

```{r}
ctrl_ser <- ctrl_par
ctrl_ser$parallel <- FALSE

t0 <- Sys.time()
cv_ser <- bhcrr_cv(x_std, y,
                   s0_seq     = s0_seq,
                   s1_seq     = s1_seq,
                   control    = ctrl_ser,
                   folds      = folds_par,           # identical folds
                   fold_inits = cv_par$fold_inits)   # identical starting points
elapsed_ser <- Sys.time() - t0

stopifnot(
  isTRUE(all.equal(cv_par$tuning,      cv_ser$tuning,      tolerance = 1e-12)),
  isTRUE(all.equal(cv_par$pooled,      cv_ser$pooled,      tolerance = 1e-12)),
  isTRUE(all.equal(cv_par$fold_scores, cv_ser$fold_scores, tolerance = 1e-12)),
  identical(cv_par$iterations,  cv_ser$iterations),
  identical(cv_par$convergence, cv_ser$convergence)
)
cat("parallel reproduces serial exactly\n")

c(parallel_secs = as.numeric(elapsed_par, units = "secs"),
  serial_secs   = as.numeric(elapsed_ser, units = "secs"),
  speedup       = as.numeric(elapsed_ser, units = "secs") /
                  as.numeric(elapsed_par, units = "secs"),
  workers       = cv_par$timing$workers)
```

Speedup is bounded by `nfolds * ncv` = 20 chains regardless of core count, and the serial comparison above reuses the cached inits while the parallel run computed them, so it slightly understates the true speedup.

### P.3 — Failures and convergence

```{r}
nrow(cv_par$errors)                          # hard failures; expect 0
sum(cv_par$tuning$n_not_converged)           # non-converged fits
sum(cv_par$tuning$n_not_converged) / prod(dim(cv_par$fold_scores))

# Where does non-convergence sit in the grid? Trimming s0 at 0.08 should have
# removed most of the region that was failing before.
conv_by_pair <- cbind(cv_par$grid[, c("pair", "s0", "s1")],
                      n_not_conv = cv_par$tuning$n_not_converged)
subset(conv_by_pair, n_not_conv > 0)

# Does the winner depend on non-converged fits?
cv_par$tuning[which.max(cv_par$tuning$score_mean), ]
```

### P.4 — Tuning result

```{r}
cv_par$best

head(cv_par$tuning[order(-cv_par$tuning$score_mean), ], 8)

# Score surface across the grid
with(cbind(cv_par$grid, score = cv_par$tuning$score_mean),
     tapply(score, list(s0 = s0, s1 = s1), identity))

theta <- mean(untuned$coefficients$Estimate != 0)
bhcrr_tune_validate(cv_par$tuning,
                    n      = nrow(x_std),
                    theta  = theta,
                    s0_seq = s0_seq,
                    s1_seq = s1_seq)
```

An `interior_max` flag here means the optimum sits on a grid edge and the grid should be widened in that direction before the result is trusted.

### P.5 — Final model at the tuned scales

Refit on the full data at `cv_par$best`, using the **same** `maxit` the CV fits used.

```{r}
final_par <- fit_ssl_psdh(
  x_std, y,
  ss          = c(cv_par$best$s0, cv_par$best$s1),
  init_method = "LASSO_bic",
  maxit       = maxit_cv,        # same budget as the CV fits
  epsilon     = 1e-04
)

final_par$conv
final_par$iterations
final_par$init_method
final_par$init_meta[c("lambda", "bic", "df")]

est <- final_par$coefficients
sum(est$Estimate != 0)
head(est[order(-abs(est$Estimate)), ], 20)
```

#### Selection against the known truth

The simulation planted its active predictors, so the final model can be scored on selection as well as on discrimination. `beta1` holds the true cause-1 effects.

```{r}
truth_active <- which(beta1 != 0)
sel_active   <- which(est$Estimate != 0)

tp <- length(intersect(sel_active, truth_active))
fp <- length(setdiff(sel_active, truth_active))
fn <- length(setdiff(truth_active, sel_active))

c(true_active = length(truth_active),
  selected    = length(sel_active),
  TP = tp, FP = fp, FN = fn,
  sensitivity = tp / length(truth_active),
  FDR         = if (length(sel_active) > 0) fp / length(sel_active) else NA_real_)

# Estimated vs true effect for every genuinely active predictor
data.frame(
  variable = est$Variable[truth_active],
  truth    = beta1[truth_active],
  estimate = est$Estimate[truth_active],
  block    = design$meta$block[truth_active]
)
```

#### Discrimination

```{r}
lp_par <- predict_from_ssl_psdh(final_par, newx = x_std, tau)
wolbers_c(y, lp_par, tau)          # IN-SAMPLE
cv_par$best$score_mean             # out-of-sample, from the CV
```

The first number is in-sample — the model was fit on all of `x_std` — so it is optimistic and not comparable to `cv_par$best$score_mean`. The gap between them is the rough size of the optimism, and the CV number is the one to report.
